### 2020-06-17 paper/main.tex
paper: Initial paper

% $Id: template.tex 11 2007-04-03 22:25:53Z jpeltier $

\documentclass{vgtc}                          % final (conference style)
%\documentclass[review]{vgtc}                 % review
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint]{vgtc}               % preprint
%\documentclass[electronic]{vgtc}             % electronic version

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Figures should be in CMYK or Grey scale format, otherwise, colour
%% shifting may occur during the printing process.

%% These few lines make a distinction between latex and pdflatex calls and they
%% bring in essential packages for graphics and font handling.
%% Note that due to the \DeclareGraphicsExtensions{} call it is no longer necessary
%% to provide the the path and extension of a graphics file:
%% \includegraphics{diamondrule} is completely sufficient.
%%
\ifpdf%                                % if we use pdflatex
  \pdfoutput=1\relax                   % create PDFs from pdfLaTeX
  \pdfcompresslevel=9                  % PDF Compression
  \pdfoptionpdfminorversion=7          % create PDF 1.7
  \ExecuteOptions{pdftex}
  \usepackage{graphicx}                % allow us to embed graphics files
  \DeclareGraphicsExtensions{.pdf,.png,.jpg,.jpeg} % for pdflatex we expect .pdf, .png, or .jpg files
\else%                                 % else we use pure latex
  \ExecuteOptions{dvips}
  \usepackage{graphicx}                % allow us to embed graphics files
  \DeclareGraphicsExtensions{.eps}     % for pure latex we expect eps files
\fi%

%% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
\graphicspath{{figures/}{pictures/}{images/}{./}} % where to search for the images

\usepackage{microtype}                 % use micro-typography (slightly more compact, better to read)
\PassOptionsToPackage{warn}{textcomp}  % to address font issues with \textrightarrow
\usepackage{textcomp}                  % use better special symbols
\usepackage{mathptmx}                  % use matching math font
\usepackage{times}                     % we use Times as the main font
\renewcommand*\ttdefault{txtt}         % a nicer typewriter font
\usepackage{cite}                      % needed to automatically sort the references
\usepackage{tabu}                      % only used for the table example
\usepackage{booktabs}                  % only used for the table example
%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

\usepackage{caption}
\captionsetup[table]{skip=10pt}
\usepackage{tabularx}
\usepackage{todonotes} %TODO: Remove after all TODO's are gone

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{Type Visualization for R}

%% This is how authors are specified in the conference style

%% Author and Affiliation (single author).
%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}}
%%\affiliation{\scriptsize Allied Widgets Research}

%% Author and Affiliation (multiple authors with single affiliations).
%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com} %
%%\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com} %
%%\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}}
%%\affiliation{\scriptsize Martha Stewart Enterprises \\ Microsoft Research}

%% Author and Affiliation (multiple authors with multiple affiliations)
\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ %
        \scriptsize Starbucks Research %
\and Ed Grimley\thanks{e-mail: ed.grimley@aol.com}\\ %
     \scriptsize Grimley Widgets, Inc. %
\and Martha Stewart\thanks{e-mail: martha.stewart@marthastewart.com}\\ %
     \parbox{1.4in}{\scriptsize \centering Martha Stewart Enterprises \\ Microsoft Research}}

%% A teaser figure can be included as follows, but is not recommended since
%% the space is now taken up by a full width abstract.
%\teaser{
%  \includegraphics[width=1.5in]{sample.eps}
%  \caption{Lookit! Lookit!}
%}

%% Abstract section.
\abstract{
  Data-driven approaches to programming language design are uncommon.
  Despite the availability of large code repositories, distilling useful
  information from programs remain difficult. Important dimensions of
  code, like run-time type data, are inscrutable without appropriate tools.
  We built {\sc TypeVis} to explore and analyze run-time types in
  the R programming language by means of visualization.
  Our system focuses on understanding and comparing the type signatures of
  function calls across the R ecosystem.
  {\sc TypeVis} uses flows to intuitively represent categorical
  and numerical aspects of type signatures.
  Insights derived from our visualization are aimed at informing
  language design decisions---specifically of a gradual type system
  currently being developed for R.

  \todo{Fill in keywords below.}
} % end of abstract

%% ACM Computing Classification System (CCS).
%% See <http://www.acm.org/about/class> for details.
%% We recommend the 2012 system <http://www.acm.org/about/class/class/2012>
%% For the 2012 system use the ``\CCScatTwelve'' which command takes four arguments.
%% The 1998 system <http://www.acm.org/about/class/class/2012> is still possible
%% For the 1998 system use the ``\CCScat'' which command takes four arguments.
%% In both cases the last two arguments (1998) or last three (2012) can be empty.

\CCScatlist{
%  \CCScatTwelve{Human-centered computing}{Visu\-al\-iza\-tion}{Visu\-al\-iza\-tion techniques}{Treemaps};%  \CCScatTwelve{Human-centered computing}{Visu\-al\-iza\-tion}{Visualization design and evaluation methods}{}
}

%\CCScatlist{
  %\CCScat{H.5.2}{User Interfaces}{User Interfaces}{Graphical user interfaces (GUI)}{};
  %\CCScat{H.5.m}{Information Interfaces and Presentation}{Miscellaneous}{}{}
%}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

%% \section{Introduction} %for journal use above \firstsection{..} instead

Programming languages commonly evolve by decree. The designer
decides that a new feature is necessary or that a past feature was
ill-conceived, and thus the language and its users move forward.
Sometimes these changes are the result of rigorous theory
or the product of community feedback. However, language design is
rarely informed by empirical data about how programs are actually
written in practice.

For a popular language, data collection is no challenge. Vast
quantities of publicly available code are hosted on open source
repositories like Github and language-specific package servers like npm.
Rather, the difficulty lies with analyzing and interpreting the data
at hand. Code is complex and highly structured. Simple
semantically-ignorant metrics like ``lines of code''
carry little significance.

To analyze one aspect of programming languages, we built {\sc TypeVis},
an interactive visualization of run-time types for the language R.
{\sc TypeVis} builds on a dataset of concrete execution traces
over the most widely used libraries in the R ecosystem.
Our visualization supports three broad tasks:
filtering the dataset down to a digestible subset,
understanding the types of function arguments and returns,
and comparing across different functions.
\todo{Explain the components of the design and how they support these tasks.}

These tasks can assist language designers during different
phases of development.
For example, exploratory data analysis can weed out designs that
are incompatible with existing programs.
\todo{Give specific motivation, i.e. gradual typing.}

Supplemental material, including the data and source code of
{\sc TypeVis}, is freely available: \todo{Fill with archival link}

%%

\section{Background and Related Work}

\todo{
  We need to take a closer look at the papers we cite.
}

%%

\section{Data}

\todo{
  Can we get a copy of Alexi's paper? This should summarize the
  dynamic analysis, what it records, and how large it is.
}

%%

\section{Task Abstraction}

\bgroup
\def\arraystretch{1.75}
\begin{table*}
  \centering
  \begin{tabularx}{\linewidth}{c|c|c|X}
    Domain Goals & Search Task & Query Task & \multicolumn{1}{c}{Abstract Task Description} \\
    \hline
    Find Function & Locate & Identify & Finding a function is a location task, where the target function is known, but its location within the package hierarchy is not. Once the desired function is found, a user must be able to identify data of interest. \\

    Determine Types & Browse & Identify & Determining types is a browse task, where the target type signature is unknown, but the location of the function of interest is known. While browsing, a user must be able to identify the frequencies of particular type signatures.\\

    Compare Signatures & Lookup & Compare & Comparing type signatures is a lookup task where both target type signatures have already been located. A user must be able to how often a signature is used compared to another, both within a function and across different functions.\\
  \end{tabularx}
  \caption{Domain goals and their mid-level and low-level task abstractions.}
  \label{tab:tasks}
\end{table*}
\egroup

\todo{Write.}

%%

\section{Design and Implementation}

\todo{Write.}

%%

\section{Discussion}

\todo{Should say this is not R specific. Point limitations.}

%%

\section{Conclusion and Future Work}

\todo{Write.}

%%
%\section{Using the Style Template}
%
%\begin{itemize}
%\item If you receive compilation errors along the lines of ``\texttt{Package ifpdf Error: Name clash, \textbackslash ifpdf is already defined}'' then please add a new line ``\texttt{\textbackslash let\textbackslash ifpdf\textbackslash relax}'' right after the ``\texttt{\textbackslash documentclass[journal]\{vgtc\}}'' call. Note that your error is due to packages you use that define ``\texttt{\textbackslash ifpdf}'' which is obsolete (the result is that \texttt{\textbackslash ifpdf} is defined twice); these packages should be changed to use ifpdf package instead.
%\item The style uses the hyperref package, thus turns references into internal links. We thus recommend to make use of the ``\texttt{\textbackslash autoref\{reference\}}'' call (instead of ``\texttt{Figure\~{}\textbackslash ref\{reference\}}'' or similar) since ``\texttt{\textbackslash autoref\{reference\}}'' turns the entire reference into an internal link, not just the number. Examples: \autoref{fig:sample} and \autoref{tab:vis_papers}.
%\item The style automatically looks for image files with the correct extension (eps for regular \LaTeX; pdf, png, and jpg for pdf\LaTeX), in a set of given subfolders (figures/, pictures/, images/). It is thus sufficient to use ``\texttt{\textbackslash includegraphics\{CypressView\}}'' (instead of ``\texttt{\textbackslash includegraphics\{pictures/CypressView.jpg\}}'').
%\item For adding hyperlinks and DOIs to the list of references, you can use ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' (instead of ``\texttt{\textbackslash bibliographystyle\{abbrv\}}''). It uses the doi and url fields in a bib\TeX\ entry and turns the entire reference into a link, giving priority to the doi. The doi can be entered with or without the ``\texttt{http://dx.doi.org/}'' url part. See the examples in the bib\TeX\ file and the bibliography at the end of this template.\\[1em]
%\textbf{Note 1:} occasionally (for some \LaTeX\ distributions) this hyper-linked bib\TeX\ style may lead to \textbf{compilation errors} (``\texttt{pdfendlink ended up in different nesting level ...}'') if a reference entry is broken across two pages (due to a bug in hyperref). In this case make sure you have the latest version of the hyperref package (i.\,e., update your \LaTeX\ installation/packages) or, alternatively, revert back to ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-narrow\}}'' (at the expense of removing hyperlinks from the bibliography) and try ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' again after some more editing.\\[1em]
%\textbf{Note 2:} the ``\texttt{-narrow}'' versions of the bibliography style use the font ``PTSansNarrow-TLF'' for typesetting the DOIs in a compact way. This font needs to be available on your \LaTeX\ system. It is part of the \href{https://www.ctan.org/pkg/paratype}{``paratype'' package}, and many distributions (such as MikTeX) have it automatically installed. If you do not have this package yet and want to use a ``\texttt{-narrow}'' bibliography style then use your \LaTeX\ system's package installer to add it. If this is not possible you can also revert to the respective bibliography styles without the ``\texttt{-narrow}'' in the file name.\\[1em]
%DVI-based processes to compile the template apparently cannot handle the different font so, by default, the template file uses the \texttt{abbrv-doi} bibliography style but the compiled PDF shows you the effect of the \texttt{abbrv-doi-hyperref-narrow} style.
%\end{itemize}
%
%\section{Bibliography Instructions}
%
%\begin{itemize}
%\item Sort all bibliographic entries alphabetically but the last name of the first author. This \LaTeX/bib\TeX\ template takes care of this sorting automatically.
%\item Merge multiple references into one; e.\,g., use \cite{Max:1995:OMF,Kitware:2003} (not \cite{Kitware:2003}\cite{Max:1995:OMF}). Within each set of multiple references, the references should be sorted in ascending order. This \LaTeX/bib\TeX\ template takes care of both the merging and the sorting automatically.
%\item Verify all data obtained from digital libraries, even ACM's DL and IEEE Xplore  etc.\ are sometimes wrong or incomplete.
%\item Do not trust bibliographic data from other services such as Mendeley.com, Google Scholar, or similar; these are even more likely to be incorrect or incomplete.
%\item Articles in journal---items to include:
%  \begin{itemize}
%  \item author names
%	\item title
%	\item journal name
%	\item year
%	\item volume
%	\item number
%	\item month of publication as variable name (i.\,e., \{jan\} for January, etc.; month ranges using \{jan \#\{/\}\# feb\} or \{jan \#\{-{}-\}\# feb\})
%  \end{itemize}
%\item use journal names in proper style: correct: ``IEEE Transactions on Visualization and Computer Graphics'', incorrect: ``Visualization and Computer Graphics, IEEE Transactions on''
%\item Papers in proceedings---items to include:
%  \begin{itemize}
%  \item author names
%	\item title
%	\item abbreviated proceedings name: e.\,g., ``Proc.\textbackslash{} CONF\_ACRONYNM'' without the year; example: ``Proc.\textbackslash{} CHI'', ``Proc.\textbackslash{} 3DUI'', ``Proc.\textbackslash{} Eurographics'', ``Proc.\textbackslash{} EuroVis''
%	\item year
%	\item publisher
%	\item town with country of publisher (the town can be abbreviated for well-known towns such as New York or Berlin)
%  \end{itemize}
%\item article/paper title convention: refrain from using curly brackets, except for acronyms/proper names/words following dashes/question marks etc.; example:
%\begin{itemize}
%	\item paper ``Marching Cubes: A High Resolution 3D Surface Construction Algorithm''
%	\item should be entered as ``\{M\}arching \{C\}ubes: A High Resolution \{3D\} Surface Construction Algorithm'' or  ``\{M\}arching \{C\}ubes: A high resolution \{3D\} surface construction algorithm''
%	\item will be typeset as ``Marching Cubes: A high resolution 3D surface construction algorithm''
%\end{itemize}
%\item for all entries
%\begin{itemize}
%	\item DOI can be entered in the DOI field as plain DOI number or as DOI url; alternative: a url in the URL field
%	\item provide full page ranges AA-{}-BB
%\end{itemize}
%\item when citing references, do not use the reference as a sentence object; e.\,g., wrong: ``In \cite{Lorensen:1987:MCA} the authors describe \dots'', correct: ``Lorensen and Cline \cite{Lorensen:1987:MCA} describe \dots''
%\end{itemize}
%
%\section{Example Section}
%
%Lorem\marginpar{\small You can use the margins for comments while editing the submission, but please remove the marginpar comments for submission.} ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
%dolore magna aliquyam erat, sed diam
%voluptua~\cite{Kitware:2003,Max:1995:OMF}. At vero eos et accusam et
%justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
%takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
%amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
%invidunt ut labore et dolore magna aliquyam erat, sed diam
%voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est.
%
%\section{Exposition}
%
%Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
%molestie consequat, vel illum dolore eu feugiat nulla facilisis at
%vero eros et accumsan et iusto odio dignissim qui blandit praesent
%luptatum zzril delenit augue duis dolore te feugait nulla
%facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
%sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
%aliquam erat volutpat~\cite{Kindlmann:1999:SAG}.
%
%\begin{equation}
%\sum_{j=1}^{z} j = \frac{z(z+1)}{2}
%\end{equation}
%
%Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
%takimata sanctus est Lorem ipsum dolor sit amet.
%
%\subsection{Lorem ipsum}
%
%Lorem ipsum dolor sit amet (see \autoref{tab:vis_papers}), consetetur sadipscing elitr, sed diam
%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
%takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
%amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
%invidunt ut labore et dolore magna aliquyam erat, sed diam
%voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum.
%
%\subsection{Mezcal Head}
%
%Lorem ipsum dolor sit amet (see \autoref{fig:sample}), consetetur sadipscing elitr, sed diam
%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
%takimata sanctus est Lorem ipsum dolor sit amet.
%
%\subsubsection{Duis Autem}
%
%Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
%takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
%amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
%invidunt ut labore et dolore magna aliquyam erat, sed diam
%voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est. Lorem
%ipsum dolor sit amet.
%
%\begin{figure}[tb]
% \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
% \includegraphics[width=\columnwidth]{paper-count-w-2015-new}
% \caption{A visualization of the 1990--2015 data from \autoref{tab:vis_papers}. The image is from \cite{Isenberg:2017:VMC} and is in the public domain.}
% \label{fig:sample}
%\end{figure}
%
%\subsubsection{Ejector Seat Reservation}
%
%Duis autem~\cite{Lorensen:1987:MCA}\footnote{The algorithm behind
%Marching Cubes \cite{Lorensen:1987:MCA} had already been
%described by Wyvill et al. \cite{Wyvill:1986:DSS} a year
%earlier.} vel eum iriure dolor in hendrerit
%in vulputate velit esse molestie consequat,\footnote{Footnotes
%appear at the bottom of the column.} vel illum dolore eu
%feugiat nulla facilisis at vero eros et accumsan et iusto odio
%dignissim qui blandit praesent luptatum zzril delenit augue duis
%dolore te feugait nulla facilisi. Lorem ipsum dolor sit amet,
%consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt
%ut laoreet dolore magna aliquam erat volutpat.
%
%
%\paragraph{Confirmed Ejector Seat Reservation}
%
%Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
%suscipit lobortis nisl ut aliquip ex ea commodo
%consequat~\cite{Nielson:1991:TAD}. Duis autem vel eum iriure dolor in
%hendrerit in vulputate velit esse molestie consequat, vel illum dolore
%eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
%dignissim qui blandit praesent luptatum zzril delenit augue duis
%dolore te feugait nulla facilisi.
%
%\paragraph{Rejected Ejector Seat Reservation}
%
%Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
%suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
%vel eum iriure dolor in hendrerit in vulputate velit esse molestie
%
%
%\section{Conclusion}
%
%Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
%takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
%amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
%invidunt ut labore et dolore magna aliquyam erat, sed diam
%voluptua. At vero eos et accusam et justo duo dolores et ea
%rebum.

%% if specified like this the section will be committed in review mode
\acknowledgments{
The authors wish to thank A, B, and C. This work was supported in part by
a grant from XYZ.}

%\bibliographystyle{abbrv}
\bibliographystyle{abbrv-doi}
%\bibliographystyle{abbrv-doi-narrow}
%\bibliographystyle{abbrv-doi-hyperref}
%\bibliographystyle{abbrv-doi-hyperref-narrow}

\bibliography{main}
\end{document}

### 2020-06-17 paper/main.tex
paper: Add images

@@
 
-\todo{Write.}
+We interviewed the primary researcher developing R's gradual type system
+to inventory relevant domain-specific tasks.
+Additionally, we sought feedback throughout the creation of {\sc TypeVis}
+to ensure it was adequately satisfying requirements.
+The domain-specific goals can be reduced to:
+
+\begin{enumerate}
+\item {\bf Find Function.} Quickly search for a particular function and identify basic information. This includes how often the function is called and which package contains it.
+\item {\bf Determine Types.} Given a specific function, determine the argument and return types of the function.
+Additionally, identify the frequency of a type signature.
+A user should be able to infer, for example, if the function is polymorphic or if it's a predicate.
+\item {\bf Compare Signatures.} Compare the occurrence of type signatures within a specific function, but also across many different ones.
+Users must be able to isolate individual signatures and see them in the context of other signatures and functions.
+\end{enumerate}
+
+These goals can be classified using an abstract task analysis framework.
+In table~\ref{tab:tasks}, we use Munzner's mid-level and low-level
+abstract task taxonomy. \todo{Cite Munzner}
 
@@
 
+\begin{figure*}
+ \centering
+ \includegraphics[width=\linewidth]{img/typevis.png}
+ \caption{The main view of {\sc TypeVis}, with type flows on the left and the package hierarchy on the right.}
+ \label{fig:typevis}
+\end{figure*}
+
+\begin{figure}[tb]
+ \centering
+ \includegraphics[width=\columnwidth]{img/no_decross.png}
+ \includegraphics[width=\columnwidth]{img/decross.png}
+ \caption{Comparison of type flows without decrossing (above) and with decrossing (below).}
+ \label{fig:decross}
+\end{figure}
+
 \todo{Write.}
+
+\subsection{Package and Function Filtering}
+
+\subsection{Types as Flows}
+
+\subsection{Technical Details}
 

### 2020-06-18 paper/main.tex
paper: Finish intro

@@
 and comparing across different functions.
-\todo{Explain the components of the design and how they support these tasks.}
+For filtering, {\sc TypeVis} uses the treemap idiom
+to convey the relative scale of different packages.
+Once a specific function is selected,
+type signatures are represented as flows over types
+where the width of the flow is proportional to frequency.
+These flows can be readily compared within the same function,
+or across different functions.
 
-These tasks can assist language designers during different
-phases of development.
+Tasks enabled by {\sc TypeVis} can assist language designers
+during many phases of development.
 For example, exploratory data analysis can weed out designs that
 are incompatible with existing programs.
-\todo{Give specific motivation, i.e. gradual typing.}
-
+In particular, the visualization is built to answer questions
+relevant to gradual type system development.
 Supplemental material, including the data and source code of
@@
 These goals can be classified using an abstract task analysis framework.
-In table~\ref{tab:tasks}, we use Munzner's mid-level and low-level
-abstract task taxonomy. \todo{Cite Munzner}
+See table~\ref{tab:tasks} for a categorization of these domain tasks
+into Munzner's mid-level and low-level abstract task taxonomy. \todo{Cite Munzner}
 
@@
 
-\subsection{Package and Function Filtering}
+\subsection{Function Filtering}
 

### 2020-06-18 paper/main.tex
paper: Begin design section

@@
 %% Author and Affiliation (multiple authors with multiple affiliations)
-\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}\\ %
+\author{Roy G. Biv\thanks{roy.g.biv@aol.com}\\ %
         \scriptsize Starbucks Research %
-\and Ed Grimley\thanks{e-mail: ed.grimley@aol.com}\\ %
+\and Ed Grimley\thanks{ed.grimley@aol.com}\\ %
      \scriptsize Grimley Widgets, Inc. %
-\and Martha Stewart\thanks{e-mail: martha.stewart@marthastewart.com}\\ %
+\and Martha Stewart\thanks{martha.stewart@marthastewart.com}\\ %
      \parbox{1.4in}{\scriptsize \centering Martha Stewart Enterprises \\ Microsoft Research}}
@@
     Compare Signatures & Lookup & Compare & Comparing type signatures is a lookup task where both target type signatures have already been located.
-A user must be able to how often a signature is used compared to another, both within a function and across different functions.\\
+A user must be able to discern how often a signature is used compared to another, both within a function and across different functions.\\
   \end{tabularx}
@@
 
+{\sc TypeVis} is an interactive web-based visualization
+that permits real-time exploration,
+even with large datasets.
+The default dataset contains over a million rows
+based on the analysis of almost five hundred
+different R packages.
+Figure~\ref{fig:typevis} displays
+the heart of {\sc TypeVis}
+in its initial configuration.
+Most aspects of the system can be tweaked
+by a user if desired.
+
+\subsection{Function Filtering}
+
+The right-hand panel of figure~\ref{fig:typevis}
+contains three elements: a search bar,
+a package treemap,
+and a function treemap.
+All three are linked.
+Selecting a package
+in the top treemap will immediately narrow the
+bottom view to only functions available in that package.
+Additionally, entering a function name in the search
+will select it in the function treemap.
+
+These different components support different modes of search.
+If a user has a specific function in mind,
+the search bar supports fuzzy autocomplete
+that can quickly select a function among thousands.
+However, if one wants to browse
+all available functions and packages,
+the treemap view is more applicable.
+Nodes in the bottom treemap convey the log-scaled
+frequency of function calls,
+and correspondingly the top treemap
+is scaled based on invocations
+of functions defined in that package.
+Log-scaling and pagination keep
+elements of the treemap legible,
+even if absolute size differences are significant.
+
+If configured, {\sc TypeVis} allows for multiple function selection.
+Both the {\tt length.unit} and {\tt valid.unit} functions
+from the {\tt grid} package are selected in figure~\ref{fig:typevis}.
+These functions are rendered simultaneously in the type flow
+panel, allowing one to make comparisons between them.
+
+\subsection{Types as Flows}
+
+Prominently featured as the center visualization,
+a Sankey diagram encodes type signatures of the selected
+functions as flows.
+Flows begin at nodes labeling a function,
+curve across each argument type in order,
+and terminate at the return type.
+Widths are proportional to how many times a function
+is called with that signature during analysis.
+Flow segments are filled with a hue-varying color scale,
+based on the type where the segment ends.
+Beneath each node label is an approximate count.
+If the node represents a function, then the
+number is how many times the function was invoked.
+If the node represents a type, then the
+number is how often a value of that type
+at that argument position was recorded during analysis.
+
+For example, in figure~\ref{fig:typevis},
+the {\tt length.unit} function is called with
+two different type signatures at run-time.
+Most frequently, the function is used with the
+signature {\tt double $\to$ integer},
+but a substantial fraction of the time it is
+used as {\tt double[] $\to$ integer}.
+Here, {\tt double[]} stands for an array of
+doubles with arbitrary dimension.
+
+Interactivity significantly augments the
+usability of the Sankey visualization.
+Hovering over a flow will highlight the
+entire path of the flow
+and will fade other functions
+into the background.
+When displaying many flows at the same time,
+highlighting becomes especially critical
+for user comprehension.
+Additional quantitative information
+about the flow is also supplied
+while hovering.
+Clicking a flow will focus on that function
+only, filtering away all others.
+
+% decrossing
+
+\subsection{Technical Details}
+
+% d3, vue
+
+% backend node, sqlite3, indexing
+
 \begin{figure*}
@@
 \end{figure}
-
-\todo{Write.}
-
-\subsection{Function Filtering}
-
-\subsection{Types as Flows}
-
-\subsection{Technical Details}
 

### 2020-06-18 paper/main.tex
paper: Finish paragraph on decrossing

@@
 
-% decrossing
+Flow layout is a critical consideration
+for a sufficiently legible visualization.
+Figure~\ref{fig:decross} exemplifies what
+happens if flow layout is not adequately handled.
+The top layout was generated without any
+flow decrossing algorithm,
+while the bottom layout does
+apply decrossing.
+Even with highlighting,
+the utility of the visualization is vastly
+reduced without some decrossing effort.
+To maintain real-time performance,
+we use two different decrossing methods.
+First, the system attempts to construct
+a layout with the globally optimal
+minimum amount of flow crossing by
+solving a mixed-integer linear program.
+Unfortunately, for even moderately
+sized graphs this can take too long to compute.
+Therefore, after about $1$ second
+the optimal algorithm is terminated
+if it has not finished, and a faster
+algorithm that only locally minimizes crossing
+between layers computes the layout.
 

### 2020-06-18 paper/main.tex
paper: Add technical details section

@@
 If a user has a specific function in mind,
-the search bar supports fuzzy autocomplete
+the search bar supports autocomplete
 that can quickly select a function among thousands.
@@
 
-% d3, vue
+Since the amount of data that a user may query is large,
+the application consists of both a frontend and a backend.
 
-% backend node, sqlite3, indexing
+On the frontend, state management, interactivity,
+and DOM manipulation is all done within the Vue framework.
+This is in contrast with many web-based visualizations
+that use D3 for DOM manipulation.
+We chose this setup because of the many different
+individual components whose state must be kept
+sychronized.
+This includes not only the visualizations, but also
+user interface elements like the autocomplete search bar.
+In our experience, Vue's model of reactivity is
+more successful at maintaining consistency across the
+entire system than using D3 and vanilla JavaScript.
+While {\sc TypeVis} does not use D3 for DOM manipulation,
+it still does use D3 for other visualization-related
+computations.
+
+Our backend is a Node.js application that serves
+JSON in a standard REST-style architecture.
+Data is stored in a SQLite database that has been
+indexed to achieve lookups as fast as possible.
 
@@
 
-\todo{Should say this is not R specific. Point limitations.}
+\todo{Point limitations.}
 
@@
 \section{Conclusion and Future Work}
+
+% more type system features (dimensionality, dispatch)
+% other languages
+% general philosophy, not just about types
 

### 2020-06-18 paper/main.tex
paper: Initial limitations and future work

@@
 
-\section{Discussion}
+\section{Limitations and Future Work}
 
-\todo{Point limitations.}
+{\sc TypeVis} suffers from a number of shortcomings
+and does not fully address the entire range of possible tasks
+a language designer may need to perform.
+
+In particular, R supports a wide range of data types
+including types that contain dimensionality information.
+For example, an R array may have type {\tt integer[3]}
+indicating that it is an array of length $3$.
+This will get compressed down into a single array
+type labeled {\tt integer[]}.
+Reducing the number of types can make the visualization
+more understandable---at the expensive of losing some
+precision in type information.
+Ideally a user would be able to interactively tune
+the level of type granularity based on their interest
+in a specific property.
+Such a feature would be especially helpful for R
+since values can come attached with relevant additional
+data such as the dynamic dispatch method
+or user-specified metadata.
+
+Filtering is an central mechanism in the system,
+but it also comes at a cost.
+If one has a specific function or several functions
+in mind, then {\sc TypeVis} provides a useful
+local view of that type information.
+However, it does not support any kind of global
+view of type information across a large amount of
+functions. When the number of flows or nodes
+becomes too great, the visualization will start
+paginating, making comparisons over large amounts
+of data impossible.
+Aggregation and summarization would be key to
+making a global view of the data feasible---that
+remains future work.
 
@@
 
-\section{Conclusion and Future Work}
+\section{Conclusion}
 

### 2020-06-19 paper/main.tex
paper: Add data and collection section

@@
 
+% R background
+
+% gradual typing background
+
+% vis papers
+
 \todo{
@@
 
-\section{Data}
+\section{Data and Collection}
 
-\todo{
-  Can we get a copy of Alexi's paper? This should summarize the
-  dynamic analysis, what it records, and how large it is.
-}
+\todo{Alexi's paper says 412, but I see more packages in our data.}
+
+The data consists of traces from 412 packages
+containing over $760,000$ lines of R code
+and $534,000$ lines of native code.
+All libraries were obtained from CRAN,
+the most popular R code repository.
+Packages hosted on CRAN must satisfy certain criteria
+to ensure they are high quality.
+Additionally, the subset of 412 packages chosen here
+all had code coverage above 65\%.
+
+Each execution trace was recorded by running
+the test, example, and vignette code of each package.
+A vignette is a form of documentation that weaves
+together prose and executable code.
+An execution trace is a list of ``raw'' type signatures
+containing, among other attributes, the function name
+and type tags for each argument and return value.
+Our dataset consists of a table of traces
+that have been processed to reduce some fields.
+Traces were generated via a dynamic analysis
+that instruments the R virtual machine with hooks
+that intercept the required data at run-time.
+
+While {\sc TypeVis} has only been run with
+data from R execution traces, the architecture is not
+limited to R.
+Any dynamically typed language with the right run-time
+hooks could generate data for the tool.
+As long as execution trace information is adheres to
+the same database schema, {\sc TypeVis} will
+work properly without any modifications.
 
@@
 indexed to achieve lookups as fast as possible.
+To avoid execessive memory consumption and speed up computations,
+clients query data from the server on an as-needed basis.
+Thus, a stable network connection is required for
+real-time interaction.
 
@@
 
-% more type system features (dimensionality, dispatch)
-% other languages
-% general philosophy, not just about types
+% other dynamic languages
+
+% other featuers not just types
+
+% general philosophy not just about types
+% vis has role to play in language design
 
@@
 
-%%
-%\section{Using the Style Template}
-%
-%\begin{itemize}
-%\item If you receive compilation errors along the lines of ``\texttt{Package ifpdf Error: Name clash, \textbackslash ifpdf is already defined}'' then please add a new line ``\texttt{\textbackslash let\textbackslash ifpdf\textbackslash relax}'' right after the ``\texttt{\textbackslash documentclass[journal]\{vgtc\}}'' call.
-Note that your error is due to packages you use that define ``\texttt{\textbackslash ifpdf}'' which is obsolete (the result is that \texttt{\textbackslash ifpdf} is defined twice); these packages should be changed to use ifpdf package instead.
-%\item The style uses the hyperref package, thus turns references into internal links.
-We thus recommend to make use of the ``\texttt{\textbackslash autoref\{reference\}}'' call (instead of ``\texttt{Figure\~{}\textbackslash ref\{reference\}}'' or similar) since ``\texttt{\textbackslash autoref\{reference\}}'' turns the entire reference into an internal link, not just the number.
-Examples: \autoref{fig:sample} and \autoref{tab:vis_papers}.
-%\item The style automatically looks for image files with the correct extension (eps for regular \LaTeX; pdf, png, and jpg for pdf\LaTeX), in a set of given subfolders (figures/, pictures/, images/).
-It is thus sufficient to use ``\texttt{\textbackslash includegraphics\{CypressView\}}'' (instead of ``\texttt{\textbackslash includegraphics\{pictures/CypressView.jpg\}}'').
-%\item For adding hyperlinks and DOIs to the list of references, you can use ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' (instead of ``\texttt{\textbackslash bibliographystyle\{abbrv\}}'').
-It uses the doi and url fields in a bib\TeX\ entry and turns the entire reference into a link, giving priority to the doi.
-The doi can be entered with or without the ``\texttt{http://dx.doi.org/}'' url part.
-See the examples in the bib\TeX\ file and the bibliography at the end of this template.\\[1em]
-%\textbf{Note 1:} occasionally (for some \LaTeX\ distributions) this hyper-linked bib\TeX\ style may lead to \textbf{compilation errors} (``\texttt{pdfendlink ended up in different nesting level ...}'') if a reference entry is broken across two pages (due to a bug in hyperref).
-In this case make sure you have the latest version of the hyperref package (i.\,e., update your \LaTeX\ installation/packages) or, alternatively, revert back to ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-narrow\}}'' (at the expense of removing hyperlinks from the bibliography) and try ``\texttt{\textbackslash bibliographystyle\{abbrv-doi-hyperref-narrow\}}'' again after some more editing.\\[1em]
-%\textbf{Note 2:} the ``\texttt{-narrow}'' versions of the bibliography style use the font ``PTSansNarrow-TLF'' for typesetting the DOIs in a compact way.
-This font needs to be available on your \LaTeX\ system.
-It is part of the \href{https://www.ctan.org/pkg/paratype}{``paratype'' package}, and many distributions (such as MikTeX) have it automatically installed.
-If you do not have this package yet and want to use a ``\texttt{-narrow}'' bibliography style then use your \LaTeX\ system's package installer to add it.
-If this is not possible you can also revert to the respective bibliography styles without the ``\texttt{-narrow}'' in the file name.\\[1em]
-%DVI-based processes to compile the template apparently cannot handle the different font so, by default, the template file uses the \texttt{abbrv-doi} bibliography style but the compiled PDF shows you the effect of the \texttt{abbrv-doi-hyperref-narrow} style.
-%\end{itemize}
-%
-%\section{Bibliography Instructions}
-%
-%\begin{itemize}
-%\item Sort all bibliographic entries alphabetically but the last name of the first author. This \LaTeX/bib\TeX\ template takes care of this sorting automatically.
-%\item Merge multiple references into one; e.\,g., use \cite{Max:1995:OMF,Kitware:2003} (not \cite{Kitware:2003}\cite{Max:1995:OMF}).
-Within each set of multiple references, the references should be sorted in ascending order.
-This \LaTeX/bib\TeX\ template takes care of both the merging and the sorting automatically.
-%\item Verify all data obtained from digital libraries, even ACM's DL and IEEE Xplore  etc.\ are sometimes wrong or incomplete.
-%\item Do not trust bibliographic data from other services such as Mendeley.com, Google Scholar, or similar; these are even more likely to be incorrect or incomplete.
-%\item Articles in journal---items to include:
-%  \begin{itemize}
-%  \item author names
-%	\item title
-%	\item journal name
-%	\item year
-%	\item volume
-%	\item number
-%	\item month of publication as variable name (i.\,e., \{jan\} for January, etc.; month ranges using \{jan \#\{/\}\# feb\} or \{jan \#\{-{}-\}\# feb\})
-%  \end{itemize}
-%\item use journal names in proper style: correct: ``IEEE Transactions on Visualization and Computer Graphics'', incorrect: ``Visualization and Computer Graphics, IEEE Transactions on''
-%\item Papers in proceedings---items to include:
-%  \begin{itemize}
-%  \item author names
-%	\item title
-%	\item abbreviated proceedings name: e.\,g., ``Proc.\textbackslash{} CONF\_ACRONYNM'' without the year; example: ``Proc.\textbackslash{} CHI'', ``Proc.\textbackslash{} 3DUI'', ``Proc.\textbackslash{} Eurographics'', ``Proc.\textbackslash{} EuroVis''
-%	\item year
-%	\item publisher
-%	\item town with country of publisher (the town can be abbreviated for well-known towns such as New York or Berlin)
-%  \end{itemize}
-%\item article/paper title convention: refrain from using curly brackets, except for acronyms/proper names/words following dashes/question marks etc.; example:
-%\begin{itemize}
-%	\item paper ``Marching Cubes: A High Resolution 3D Surface Construction Algorithm''
-%	\item should be entered as ``\{M\}arching \{C\}ubes: A High Resolution \{3D\} Surface Construction Algorithm'' or  ``\{M\}arching \{C\}ubes: A high resolution \{3D\} surface construction algorithm''
-%	\item will be typeset as ``Marching Cubes: A high resolution 3D surface construction algorithm''
-%\end{itemize}
-%\item for all entries
-%\begin{itemize}
-%	\item DOI can be entered in the DOI field as plain DOI number or as DOI url; alternative: a url in the URL field
-%	\item provide full page ranges AA-{}-BB
-%\end{itemize}
-%\item when citing references, do not use the reference as a sentence object; e.\,g., wrong: ``In \cite{Lorensen:1987:MCA} the authors describe \dots'', correct: ``Lorensen and Cline \cite{Lorensen:1987:MCA} describe \dots''
-%\end{itemize}
-%
-%\section{Example Section}
-%
-%Lorem\marginpar{\small You can use the margins for comments while editing the submission, but please remove the marginpar comments for submission.} ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-%dolore magna aliquyam erat, sed diam
-%voluptua~\cite{Kitware:2003,Max:1995:OMF}. At vero eos et accusam et
-%justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-%takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-%amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-%invidunt ut labore et dolore magna aliquyam erat, sed diam
-%voluptua. At vero eos et accusam et justo duo dolores et ea
-%rebum. Stet clita kasd gubergren, no sea takimata sanctus est.
-%
-%\section{Exposition}
-%
-%Duis autem vel eum iriure dolor in hendrerit in vulputate velit esse
-%molestie consequat, vel illum dolore eu feugiat nulla facilisis at
-%vero eros et accumsan et iusto odio dignissim qui blandit praesent
-%luptatum zzril delenit augue duis dolore te feugait nulla
-%facilisi. Lorem ipsum dolor sit amet, consectetuer adipiscing elit,
-%sed diam nonummy nibh euismod tincidunt ut laoreet dolore magna
-%aliquam erat volutpat~\cite{Kindlmann:1999:SAG}.
-%
-%\begin{equation}
-%\sum_{j=1}^{z} j = \frac{z(z+1)}{2}
-%\end{equation}
-%
-%Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-%takimata sanctus est Lorem ipsum dolor sit amet.
-%
-%\subsection{Lorem ipsum}
-%
-%Lorem ipsum dolor sit amet (see \autoref{tab:vis_papers}), consetetur sadipscing elitr, sed diam
-%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-%takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-%amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-%invidunt ut labore et dolore magna aliquyam erat, sed diam
-%voluptua. At vero eos et accusam et justo duo dolores et ea
-%rebum.
-%
-%\subsection{Mezcal Head}
-%
-%Lorem ipsum dolor sit amet (see \autoref{fig:sample}), consetetur sadipscing elitr, sed diam
-%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-%takimata sanctus est Lorem ipsum dolor sit amet.
-%
-%\subsubsection{Duis Autem}
-%
-%Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-%takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-%amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-%invidunt ut labore et dolore magna aliquyam erat, sed diam
-%voluptua. At vero eos et accusam et justo duo dolores et ea
-%rebum. Stet clita kasd gubergren, no sea takimata sanctus est. Lorem
-%ipsum dolor sit amet.
-%
-%\begin{figure}[tb]
-% \centering % avoid the use of \begin{center}...\end{center} and use \centering instead (more compact)
-% \includegraphics[width=\columnwidth]{paper-count-w-2015-new}
-% \caption{A visualization of the 1990--2015 data from \autoref{tab:vis_papers}. The image is from \cite{Isenberg:2017:VMC} and is in the public domain.}
-% \label{fig:sample}
-%\end{figure}
-%
-%\subsubsection{Ejector Seat Reservation}
-%
-%Duis autem~\cite{Lorensen:1987:MCA}\footnote{The algorithm behind
-%Marching Cubes \cite{Lorensen:1987:MCA} had already been
-%described by Wyvill et al. \cite{Wyvill:1986:DSS} a year
-%earlier.} vel eum iriure dolor in hendrerit
-%in vulputate velit esse molestie consequat,\footnote{Footnotes
-%appear at the bottom of the column.} vel illum dolore eu
-%feugiat nulla facilisis at vero eros et accumsan et iusto odio
-%dignissim qui blandit praesent luptatum zzril delenit augue duis
-%dolore te feugait nulla facilisi. Lorem ipsum dolor sit amet,
-%consectetuer adipiscing elit, sed diam nonummy nibh euismod tincidunt
-%ut laoreet dolore magna aliquam erat volutpat.
-%
-%
-%\paragraph{Confirmed Ejector Seat Reservation}
-%
-%Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-%suscipit lobortis nisl ut aliquip ex ea commodo
-%consequat~\cite{Nielson:1991:TAD}. Duis autem vel eum iriure dolor in
-%hendrerit in vulputate velit esse molestie consequat, vel illum dolore
-%eu feugiat nulla facilisis at vero eros et accumsan et iusto odio
-%dignissim qui blandit praesent luptatum zzril delenit augue duis
-%dolore te feugait nulla facilisi.
-%
-%\paragraph{Rejected Ejector Seat Reservation}
-%
-%Ut wisi enim ad minim veniam, quis nostrud exerci tation ullamcorper
-%suscipit lobortis nisl ut aliquip ex ea commodo consequat. Duis autem
-%vel eum iriure dolor in hendrerit in vulputate velit esse molestie
-%
-%
-%\section{Conclusion}
-%
-%Lorem ipsum dolor sit amet, consetetur sadipscing elitr, sed diam
-%nonumy eirmod tempor invidunt ut labore et dolore magna aliquyam erat,
-%sed diam voluptua. At vero eos et accusam et justo duo dolores et ea
-%rebum. Stet clita kasd gubergren, no sea takimata sanctus est Lorem
-%ipsum dolor sit amet. Lorem ipsum dolor sit amet, consetetur
-%sadipscing elitr, sed diam nonumy eirmod tempor invidunt ut labore et
-%dolore magna aliquyam erat, sed diam voluptua. At vero eos et accusam
-%et justo duo dolores et ea rebum. Stet clita kasd gubergren, no sea
-%takimata sanctus est Lorem ipsum dolor sit amet. Lorem ipsum dolor sit
-%amet, consetetur sadipscing elitr, sed diam nonumy eirmod tempor
-%invidunt ut labore et dolore magna aliquyam erat, sed diam
-%voluptua. At vero eos et accusam et justo duo dolores et ea
-%rebum.
-
-%% if specified like this the section will be committed in review mode
 \acknowledgments{
   The authors wish to thank A, B, and C. This work was supported in part by
-a grant from XYZ.}
+  a grant from XYZ.
+}
 
-%\bibliographystyle{abbrv}
 \bibliographystyle{abbrv-doi}
-%\bibliographystyle{abbrv-doi-narrow}
-%\bibliographystyle{abbrv-doi-hyperref}
-%\bibliographystyle{abbrv-doi-hyperref-narrow}
 

### 2020-06-19 paper/main.tex
paper: Background

@@
 
-% R background
+R is a dynamically typed programming language widely used for
+statistical computing. Unlike statically typed languages,
+R programs do not contain type annotations
+and never pass through a typechecker.
+At run-time,
+functions may called with values of arbitrary type,
+and may return values of arbitrary type.\footnote{
+  We use the word ``type'' to mean what others may call a ``type tag.''
+  This is a run-time concept---we do not assume a static type system
+  because there is none. Indeed, the goal is to help design one!
+  A ``type signature'' in this context is one particular configuration
+  of argument and return types a function is called with at run-time.
+}
 
-% gradual typing background
+For statically typed languages, the type system is usually
+an inseparable part of the language. Since they are designed
+at the same time, the type system precedes any code written
+in the language.
+A gradual type system, by contrast,
+permits the integration of static and dynamic typing.
+Such a type system is capable of seamlessly mixing typed and untyped regions of code.
+Often, a gradual type system is designed atop
+an existing dynamically typed language.
+Since many R programs already exist,
+a successful gradual type system must be capable of
+typechecking idiomatic R code.
+{\sc TypeVis} aims to clarify how existing R programs
+use types to assist in this task.
+While this is our motivation, the visualization's use
+is not necessarily limited to this application.
+Anyone who has an interest in how the real R programs use types
+may find it useful.
 
-% vis papers
-
-\todo{
-  We need to take a closer look at the papers we cite.
-}
+\todo{Prior work}
 
@@
 \section{Design and Implementation}
+
+\begin{figure*}
+ \centering
+ \includegraphics[width=\linewidth]{img/typevis.png}
+ \caption{The main view of {\sc TypeVis}, with type flows on the left and the package hierarchy on the right.
+Here, the {\tt length.unit} and {\tt valid.unit} functions from the {\tt grid} package are both selected and displayed simultaneously.}
+ \label{fig:typevis}
+\end{figure*}
 
@@
 
-\begin{figure*}
- \centering
- \includegraphics[width=\linewidth]{img/typevis.png}
- \caption{The main view of {\sc TypeVis}, with type flows on the left and the package hierarchy on the right.}
- \label{fig:typevis}
-\end{figure*}
-
 \begin{figure}[tb]

### 2020-06-20 paper/main.tex
paper: Finish up

@@
 
-%% Paper title.
-
 \title{Type Visualization for R}
 
-%% This is how authors are specified in the conference style
+\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com} %
+\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com} %
+\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}}
+\affiliation{\scriptsize Northeastern University}
 
-%% Author and Affiliation (single author).
-%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com}}
-%%\affiliation{\scriptsize Allied Widgets Research}
-
-%% Author and Affiliation (multiple authors with single affiliations).
-%%\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com} %
-%%\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com} %
-%%\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}}
-%%\affiliation{\scriptsize Martha Stewart Enterprises \\ Microsoft Research}
-
-%% Author and Affiliation (multiple authors with multiple affiliations)
-\author{Roy G. Biv\thanks{roy.g.biv@aol.com}\\ %
-        \scriptsize Starbucks Research %
-\and Ed Grimley\thanks{ed.grimley@aol.com}\\ %
-     \scriptsize Grimley Widgets, Inc. %
-\and Martha Stewart\thanks{martha.stewart@marthastewart.com}\\ %
-     \parbox{1.4in}{\scriptsize \centering Martha Stewart Enterprises \\ Microsoft Research}}
-
-%% A teaser figure can be included as follows, but is not recommended since
-%% the space is now taken up by a full width abstract.
-%\teaser{
-%  \includegraphics[width=1.5in]{sample.eps}
-%  \caption{Lookit! Lookit!}
-%}
-
-%% Abstract section.
 \abstract{
@@
   \todo{Fill in keywords below.}
-} % end of abstract
+}
 
@@
 
-\section{Background and Related Work}
+\section{Background and Prior Work}
 
@@
 
-\todo{Prior work}
+We are not aware of any previous work that applies
+visualization techniques for the purposes of programming language design.
+However, there are existing publications on understanding
+software systems through visualization.
+Bohnet visualize massive program execution traces
+to assist developers in code comprehension of large software systems.
+Their tool applies pruning and summarization to reduce trace size.
+The visualization focuses on call topology,
+describing the structure of function invocations,
+rather than types.
+Telea also develop a methodology for displaying call structure.
+However, they visualize call graphs of systems in their entirety,
+rather than individual execution traces.
+Our visualization can be thought of as a hybrid of the two,
+investigating the system as a whole,
+but doing so via execution traces.
+In addition to standard call graphs,
+LaToza adorn nodes and edges with extra information,
+allowing the user gain further understanding of the system
+under examination.
+This includes indicating methods that come from the same class---a
+basic form of type information.
 
 %%
+
+\bgroup
+\def\arraystretch{1.75}
+\begin{table*}
+  \centering
+  \begin{tabularx}{\linewidth}{c|c|c|X}
+    Domain Goals & Search Task & Query Task & \multicolumn{1}{c}{Abstract Task Description} \\
+    \hline
+    Find Function & Locate & Identify & Finding a function is a location task, where the target function is known, but its location within the package hierarchy is not.
+Once the desired function is found, a user must be able to identify data of interest.
+\\
+
+    Determine Types & Browse & Identify & Determining types is a browse task, where the target type signature is unknown, but the location of the function of interest is known.
+While browsing, a user must be able to identify the frequencies of particular type signatures.\\
+
+    Compare Signatures & Lookup & Compare & Comparing type signatures is a lookup task where both target type signatures have already been located.
+A user must be able to discern how often a signature is used compared to another, both within a function and across different functions.\\
+  \end{tabularx}
+  \caption{Domain goals and their mid-level and low-level task abstractions.}
+  \label{tab:tasks}
+\end{table*}
+\egroup
 
@@
 \section{Task Abstraction}
-
-\bgroup
-\def\arraystretch{1.75}
-\begin{table*}
-  \centering
-  \begin{tabularx}{\linewidth}{c|c|c|X}
-    Domain Goals & Search Task & Query Task & \multicolumn{1}{c}{Abstract Task Description} \\
-    \hline
-    Find Function & Locate & Identify & Finding a function is a location task, where the target function is known, but its location within the package hierarchy is not.
-Once the desired function is found, a user must be able to identify data of interest.
-\\
-
-    Determine Types & Browse & Identify & Determining types is a browse task, where the target type signature is unknown, but the location of the function of interest is known.
-While browsing, a user must be able to identify the frequencies of particular type signatures.\\
-
-    Compare Signatures & Lookup & Compare & Comparing type signatures is a lookup task where both target type signatures have already been located.
-A user must be able to discern how often a signature is used compared to another, both within a function and across different functions.\\
-  \end{tabularx}
-  \caption{Domain goals and their mid-level and low-level task abstractions.}
-  \label{tab:tasks}
-\end{table*}
-\egroup
 
@@
 
+Treemaps, introduced by Shneiderman,
+are often used for displaying hierarchical data.
+Package systems in many languages,
+like Java,
+contain deeply nested namespace hierarchies.
+This idiom is appropriate for navigation of such
+hierarchies.
+R packages are not structured deeply
+so the treemaps in our data are only one
+level deep.
+However, if the dataset was more deeply nested,
+{\sc TypeVis} can be configured to
+handle that.
+
 If configured, {\sc TypeVis} allows for multiple function selection.
@@
 Prominently featured as the center visualization,
-a Sankey diagram encodes type signatures of the selected
+a parallel sets diagram encodes type signatures of the selected
 functions as flows.
+Parallel sets,
+proposed by Kosara,
+are ideally suited for type signatures.
+Each data point is can be thought of as
+a tuple $(f, \tau_1, \ldots, \tau_n)$
+corresponding to the function name,
+argument types,
+and return type.
+Every component is categorical,
+and we are interested in visualizing frequencies of these points.
+This is the exact use-case for parallel sets.
+
 Flows begin at nodes labeling a function,
@@
 Interactivity significantly augments the
-usability of the Sankey visualization.
+usability of the parallel sets visualization.
 Hovering over a flow will highlight the
@@
 
-% other dynamic languages
+We have presented {\sc TypeVis}, an interactive visualization
+of run-time type information for R.
+It supports real-time exploration of a dataset of
+execution traces from a large corpus of R packages.
+To narrow the data under consideration,
+{\sc TypeVis} provides facilities for filtering
+and searching.
+Type signatures are represented as flows travesing across
+each argument type and return type in sequence.
+Flow widths are proportional to how often
+that type signature was seen during the dynamic analysis.
+Interactions provide additional details and
+assist users in comprehension and navigation.
 
-% other featuers not just types
+Our tool is aimed at elucidating how R programmers
+use the language in practice.
+Specifically, the data and visualization will be used to
+inform the design of a gradual type system for R.
+However, {\sc TypeVis} is not tied to the R ecosystem
+and is suitable to use for any dynamically typed language
+with a similar dataset.
+Additionally, the visualization could yield insights
+useful for other purposes than gradual typing.
 
-% general philosophy not just about types
-% vis has role to play in language design
+We hope that, in the future, programming language designers
+will collect and use empirical data about how their languages
+are used in practice.
+Such data can be used to make informed decisions about
+language design.
+Visualization will be unavoidable in these scenarios
+since semantically-meaning information about programs
+is rich in structure.
+Generic, canned visualizations will not be sufficient.
+{\sc TypeVis} is just the first step in this direction.
 
-\todo{Write.}
-
+\todo{Finish these.}
 \acknowledgments{
-  The authors wish to thank A, B, and C. This work was supported in part by
-  a grant from XYZ.
+  The authors wish to thank the following people:
+  students of CS 7250 for useful design feedback,
+  contributors to all the open source projects that {\sc TypeVis} depends on.
+  This work was supported in part by a Northeastern Graduate Fellowship.
 }

### 2020-06-20 paper/main.tex
paper: Cites

@@
   currently being developed for R.
-
-  \todo{Fill in keywords below.}
 }
@@
 \CCScatlist{
-%  \CCScatTwelve{Human-centered computing}{Visu\-al\-iza\-tion}{Visu\-al\-iza\-tion techniques}{Treemaps};%  \CCScatTwelve{Human-centered computing}{Visu\-al\-iza\-tion}{Visualization design and evaluation methods}{}
+  \CCScatTwelve{Human-centered computing}{Visualization}{Visualization systems and tools}{}
 }
-
-%\CCScatlist{
-  %\CCScat{H.5.2}{User Interfaces}{User Interfaces}{Graphical user interfaces (GUI)}{};
-  %\CCScat{H.5.m}{Information Interfaces and Presentation}{Miscellaneous}{}{}
-%}
 
@@
 
-%% The ``\maketitle'' command must be the first command after the
-%% ``\begin{document}'' command. It prepares and prints the title block.
-
-%% the only exception to this rule is the \firstsection command
 \firstsection{Introduction}
@@
 \maketitle
-
-%% \section{Introduction} %for journal use above \firstsection{..} instead
 
@@
 software systems through visualization.
-Bohnet visualize massive program execution traces
+Bohnet et al.~\cite{bohnet:2009} visualize massive program execution traces
 to assist developers in code comprehension of large software systems.
@@
 rather than types.
-Telea also develop a methodology for displaying call structure.
+Telea et al.~\cite{telea:2009} also develop a methodology for displaying call structure.
 However, they visualize call graphs of systems in their entirety,
@@
 In addition to standard call graphs,
-LaToza adorn nodes and edges with extra information,
+LaToza and Myers~\cite{latoza:2011} adorn nodes and edges with extra information,
 allowing the user gain further understanding of the system
@@
 See table~\ref{tab:tasks} for a categorization of these domain tasks
-into Munzner's mid-level and low-level abstract task taxonomy. \todo{Cite Munzner}
+the mid-level and low-level abstract task taxonomy
+established by Brehmer and Munzner~\cite{brehmer:2013}.
 
@@
 
-Treemaps, introduced by Shneiderman,
-are often used for displaying hierarchical data.
+Treemaps are often used for displaying hierarchical data~\cite{shneiderman:1992}.
 Package systems in many languages,
@@
 Parallel sets,
-proposed by Kosara,
+proposed by Kosara et al.~\cite{kosara:2006},
 are ideally suited for type signatures.
@@
 
-\todo{Finish these.}
 \acknowledgments{
-  The authors wish to thank the following people:
-  students of CS 7250 for useful design feedback,
-  contributors to all the open source projects that {\sc TypeVis} depends on.
+  The authors wish to thank the students of CS 7250 for useful design feedback,
+  and contributors to all the open source projects that {\sc TypeVis} depends on.
   This work was supported in part by a Northeastern Graduate Fellowship.

### 2020-06-20 paper/main.tex
paper: Typos

@@
 
-\author{Roy G. Biv\thanks{e-mail: roy.g.biv@aol.com} %
-\and Ed Grimley\thanks{e-mail:ed.grimley@aol.com} %
-\and Martha Stewart\thanks{e-mail:martha.stewart@marthastewart.com}}
+\author{Cameron Moy\thanks{camoy@ccs.neu.edu}
+\and Julia Belyakova\thanks{belyakovay@ccs.neu.edu}}
 \affiliation{\scriptsize Northeastern University}
@@
 individual components whose state must be kept
-sychronized.
+synchronized.
 This includes not only the visualizations, but also
@@
 indexed to achieve lookups as fast as possible.
-To avoid execessive memory consumption and speed up computations,
+To avoid excessive memory consumption and speed up computations,
 clients query data from the server on an as-needed basis.
@@
 and searching.
-Type signatures are represented as flows travesing across
+Type signatures are represented as flows traversing across
 each argument type and return type in sequence.

### 2020-06-21 paper/main.tex
paper: Add Julia's edits

@@
 
+\usepackage{xspace}
+
 \usepackage{caption}
@@
 \usepackage{todonotes} %TODO: Remove after all TODO's are gone
+
+%% ===================================== Commands
+\newcommand{\code}[1]{\texttt{\small{#1}}}
+\newcommand{\typevis}{{\sc TypeVis}\xspace}
 
@@
   code, like run-time type data, are inscrutable without appropriate tools.
-  We built {\sc TypeVis} to explore and analyze run-time types in
+  We built \typevis to explore and analyze run-time types in
   the R programming language by means of visualization.
@@
   function calls across the R ecosystem.
-  {\sc TypeVis} uses flows to intuitively represent categorical
+  \typevis uses flows to intuitively represent categorical
   and numerical aspects of type signatures.
@@
 
-Programming languages commonly evolve by decree. The designer
+Programming languages commonly evolve by decree. Often, the designer
 decides that a new feature is necessary or that a past feature was
-ill-conceived, and thus the language and its users move forward.
-Sometimes these changes are the result of rigorous theory
-or the product of community feedback. However, language design is
-rarely informed by empirical data about how programs are actually
-written in practice.
+ill-conceived. Thus, the language moves forward,
+forcing its users to adapt to the changes.
+Sometimes, these changes are the result of rigorous theory
+or the product of community feedback. However, rarely is language design
+informed by empirical data about how programs are \emph{actually
+written} in practice.
 
-For a popular language, data collection is no challenge. Vast
+%For a popular language, data collection is no challenge. Vast
+Thanks to the large amount of open source code, collecting data about
+the usage of a popular programming language is feasible. Vast
 quantities of publicly available code are hosted on open source
-repositories like Github and language-specific package servers like npm.
-Rather, the difficulty lies with analyzing and interpreting the data
-at hand. Code is complex and highly structured. Simple
-semantically-ignorant metrics like ``lines of code''
-carry little significance.
+repositories such as Github.
+Furthermore, if there is a concern about the quality of such data
+(for example, code duplication~\cite{lopes:2017}),
+code can be obtained from curated
+language-specific package repositories such as npm or CRAN.
 
-To analyze one aspect of programming languages, we built {\sc TypeVis},
-an interactive visualization of run-time types for the language R.
-{\sc TypeVis} builds on a dataset of concrete execution traces
-over the most widely used libraries in the R ecosystem.
-Our visualization supports three broad tasks:
-filtering the dataset down to a digestible subset,
-understanding the types of function arguments and returns,
-and comparing across different functions.
-For filtering, {\sc TypeVis} uses the treemap idiom
-to convey the relative scale of different packages.
-Once a specific function is selected,
-type signatures are represented as flows over types
+To inform programming language design, all the collected data
+needs to be analyzed and interpreted.
+As code is complex and highly structured, simple semantically-ignorant
+metrics like ``lines of code'' carry little significance.
+Therefore, programming languages researchers often employ static
+and dynamic analyses to gather information about specific aspects
+of programs in the dataset, for example, the use of a language feature.
+And even then it may be hard to make sense of the data,
+especially if the dataset is large.
+%Rather, the difficulty lies with analyzing and interpreting the data
+%at hand. Code is complex and highly structured. Simple
+
+To analyze one aspect of programming languages---\emph{the use of types}---we
+built \typevis, an interactive visualization of run-time type
+signatures of function calls.
+%an interactive visualization of run-time types for the language R.
+\typevis is instantiated for the language R:
+it builds on a dataset of concrete execution traces over unit tests and vignettes
+of the most widely used libraries in the R ecosystem.
+For example, for the function call \code{!(FALSE)},
+the recorded type signature would be \code{logical->logical},
+meaning that function \code{!} took a logical value and returned a logical value.
+
+The visualization supports three broad tasks:
+\begin{itemize}
+  \item filtering the dataset down to an amenable subset;
+  \item understanding the types of function arguments and returns;
+  \item comparing types across different functions.
+\end{itemize}
+%For filtering, \typevis uses the treemap idiom
+%to convey the relative scale of different packages.
+For exploring and filtering the dataset, \typevis relies on the treemap idiom
+to represent R packages and functions, with the block sizes conveying
+the relative popularity in the dataset.
+In addition to the treemaps, a search box can be used to directly filter
+by functions.
+Once a specific subset of one or more functions is selected,
+their type signatures are represented as flows over types
 where the width of the flow is proportional to frequency.
@@
 
-Tasks enabled by {\sc TypeVis} can assist language designers
-during many phases of development.
-For example, exploratory data analysis can weed out designs that
-are incompatible with existing programs.
+Tasks enabled by \typevis can assist language designers
+during multiple phases of development.
+For example, exploratory data analysis can identify unexpected uses
+of specific functions, or weed out language designs
+that are incompatible with existing programs.
 In particular, the visualization is built to answer questions
-relevant to gradual type system development.
+relevant to the development of a gradual type system for R.
+However, the visualization is not limited to R and can be used to
+analyze type usage in any programming language where type information
+is available at run-time.
 Supplemental material, including the data and source code of
-{\sc TypeVis}, is freely available: \todo{Fill with archival link}
+\typevis, is freely available: \todo{Fill with archival link}
 
@@
 At run-time,
-functions may called with values of arbitrary type,
+functions may be called with values of arbitrary type,
 and may return values of arbitrary type.\footnote{
@@
 
+Being a dynamically typed language most actively used in a single domain,
+R is a good target for a gradual type system.
 For statically typed languages, the type system is usually
 an inseparable part of the language. Since they are designed
-at the same time, the type system precedes any code written
+at the same time, the (static) type system precedes any code written
 in the language.
@@
 typechecking idiomatic R code.
-{\sc TypeVis} aims to clarify how existing R programs
+\typevis aims to clarify how existing R programs
 use types to assist in this task.
@@
 
-The data consists of traces from 412 packages
+The data is provided by Alexi Turcotte:
+it consists of traces from 412 packages
 containing over $760,000$ lines of R code
@@
 Additionally, the subset of 412 packages chosen here
-all had code coverage above 65\%.
+all had code coverage above 65\%.\todo{I don't understand what subset? Should we remove this line?}
 
-Each execution trace was recorded by running
-the test, example, and vignette code of each package.
+Each execution trace was recorded by a dynamic analysis
+running the test, example, and vignette code of each package.
 A vignette is a form of documentation that weaves
 together prose and executable code.
-An execution trace is a list of ``raw'' type signatures
+An execution trace is a list of ``raw'' function call signatures
 containing, among other attributes, the function name
 and type tags for each argument and return value.
-Our dataset consists of a table of traces
-that have been processed to reduce some fields.
-Traces were generated via a dynamic analysis
-that instruments the R virtual machine with hooks
-that intercept the required data at run-time.
+The type tags are assigned by the dynamic analysis and
+refine over actual type tags in R; for example, the analysis
+assigns scalar tag \code{double} to a singleton vector of double values
+because R does not have native support for scalars.
+Our dataset consists of a table of reduced execution traces:
+because \typevis is designed for the analysis of types, not all information
+in the traces is relevant.
+%that have been processed to reduce some fields.
+%Traces were generated via a dynamic analysis
+%that instruments the R virtual machine with hooks
+%that intercept the required data at run-time.
 
-While {\sc TypeVis} has only been run with
+While \typevis has only been instantiated with
 data from R execution traces, the architecture is not
 limited to R.
-Any dynamically typed language with the right run-time
-hooks could generate data for the tool.
-As long as execution trace information is adheres to
-the same database schema, {\sc TypeVis} will
+Any dynamically typed language %with the right run-time
+%hooks could generate data for the tool.
+(or a statically typed language with run-time tag information)
+could be instrumented to generate data for the tool.
+As long as execution trace information adheres to
+the same database schema, \typevis will
 work properly without any modifications.
@@
 to inventory relevant domain-specific tasks.
-Additionally, we sought feedback throughout the creation of {\sc TypeVis}
+Additionally, we sought feedback throughout the creation of \typevis
 to ensure it was adequately satisfying requirements.
@@
  \includegraphics[width=\linewidth]{img/typevis.png}
- \caption{The main view of {\sc TypeVis}, with type flows on the left and the package hierarchy on the right.
+ \caption{The main view of \typevis, with type flows on the left and the package hierarchy on the right.
 Here, the {\tt length.unit} and {\tt valid.unit} functions from the {\tt grid} package are both selected and displayed simultaneously.}
@@
 
-{\sc TypeVis} is an interactive web-based visualization
+\typevis is an interactive web-based visualization
 that permits real-time exploration,
@@
 Figure~\ref{fig:typevis} displays
-the heart of {\sc TypeVis}
+the heart of \typevis
 in its initial configuration.
@@
 However, if the dataset was more deeply nested,
-{\sc TypeVis} can be configured to
+\typevis can be configured to
 handle that.
 
-If configured, {\sc TypeVis} allows for multiple function selection.
+If configured, \typevis allows for multiple function selection.
 Both the {\tt length.unit} and {\tt valid.unit} functions
@@
 entire system than using D3 and vanilla JavaScript.
-While {\sc TypeVis} does not use D3 for DOM manipulation,
+While \typevis does not use D3 for DOM manipulation,
 it still does use D3 for other visualization-related
@@
 
-{\sc TypeVis} suffers from a number of shortcomings
+\typevis suffers from a number of shortcomings
 and does not fully address the entire range of possible tasks
@@
 If one has a specific function or several functions
-in mind, then {\sc TypeVis} provides a useful
+in mind, then \typevis provides a useful
 local view of that type information.
@@
 
-We have presented {\sc TypeVis}, an interactive visualization
+We have presented \typevis, an interactive visualization
 of run-time type information for R.
@@
 To narrow the data under consideration,
-{\sc TypeVis} provides facilities for filtering
+\typevis provides facilities for filtering
 and searching.
@@
 inform the design of a gradual type system for R.
-However, {\sc TypeVis} is not tied to the R ecosystem
+However, \typevis is not tied to the R ecosystem
 and is suitable to use for any dynamically typed language
@@
 Generic, canned visualizations will not be sufficient.
-{\sc TypeVis} is just the first step in this direction.
+\typevis is just the first step in this direction.
 
@@
   The authors wish to thank the students of CS 7250 for useful design feedback,
-  and contributors to all the open source projects that {\sc TypeVis} depends on.
+  and contributors to all the open source projects that \typevis depends on.
   This work was supported in part by a Northeastern Graduate Fellowship.

### 2020-06-21 paper/main.tex
paper: My edits to abstract and intro

@@
   Despite the availability of large code repositories, distilling useful
-  information from programs remain difficult. Important dimensions of
-  code, like run-time type data, are inscrutable without appropriate tools.
-  We built \typevis to explore and analyze run-time types in
-  the R programming language by means of visualization.
-  Our system focuses on understanding and comparing the type signatures of
-  function calls across the R ecosystem.
-  \typevis uses flows to intuitively represent categorical
+  information from programs remains difficult. Important dimensions,
+  like run-time type data, are inscrutable without the appropriate tools.
+  \typevis is an interactive visualization for exploring and analyzing
+  type information from program execution traces---aiding
+  user understanding of function type signatures across many executions.
+  The system uses flows to intuitively represent categorical
   and numerical aspects of type signatures.
   Insights derived from our visualization are aimed at informing
-  language design decisions---specifically of a gradual type system
-  currently being developed for R.
+  language design decisions---specifically of a new gradual type system
+  currently being developed for the R programming language.
 }
@@
 
-Programming languages commonly evolve by decree. Often, the designer
-decides that a new feature is necessary or that a past feature was
-ill-conceived. Thus, the language moves forward,
+Programming languages commonly evolve by decree. Often, the language designer
+decides that a new feature is necessary, or that a past feature was
+ill-conceived. Thus, the language moves forward---
 forcing its users to adapt to the changes.
@@
 or the product of community feedback. However, rarely is language design
-informed by empirical data about how programs are \emph{actually
-written} in practice.
+informed by empirical data on how programs are \emph{actually
+written} in practice. Why so?
 
 %For a popular language, data collection is no challenge. Vast
-Thanks to the large amount of open source code, collecting data about
+Thanks to the prevalence of open source code, collecting data on
 the usage of a popular programming language is feasible. Vast
-quantities of publicly available code are hosted on open source
-repositories such as Github.
-Furthermore, if there is a concern about the quality of such data
-(for example, code duplication~\cite{lopes:2017}),
+quantities of publicly available code are hosted on repositories such as Github.
+Furthermore, if there is concern about the quality of this data,
+such as code duplication~\cite{lopes:2017},
 code can be obtained from curated
-language-specific package repositories such as npm or CRAN.
+language-specific package repositories like npm or CRAN.
 
-To inform programming language design, all the collected data
-needs to be analyzed and interpreted.
-As code is complex and highly structured, simple semantically-ignorant
-metrics like ``lines of code'' carry little significance.
-Therefore, programming languages researchers often employ static
-and dynamic analyses to gather information about specific aspects
-of programs in the dataset, for example, the use of a language feature.
-And even then it may be hard to make sense of the data,
-especially if the dataset is large.
+To inform programming language design,
+the collected data needs to be analyzed and interpreted.
+Programs are complex and highly structured,
+so simple semantically-ignorant metrics like ``lines of code'' carry little significance.
+Therefore, researchers often employ static and dynamic analyses
+to gather information about specific aspects of programs.
+Even then, it may be difficult to make sense of the
+results of these analyses,
+especially if the data set is large.
 %Rather, the difficulty lies with analyzing and interpreting the data
@@
 
-To analyze one aspect of programming languages---\emph{the use of types}---we
-built \typevis, an interactive visualization of run-time type
-signatures of function calls.
+To analyze just one aspect of programming languages, \emph{type usage}, we
+built \typevis, an interactive visualization of run-time type signatures.
 %an interactive visualization of run-time types for the language R.
-\typevis is instantiated for the language R:
-it builds on a dataset of concrete execution traces over unit tests and vignettes
+\typevis was created with the R programming language in mind,
+and builds on a data set of execution traces
+mined from the unit tests and vignettes
 of the most widely used libraries in the R ecosystem.
-For example, for the function call \code{!(FALSE)},
-the recorded type signature would be \code{logical->logical},
-meaning that function \code{!} took a logical value and returned a logical value.
 
-The visualization supports three broad tasks:
+The tool supports three broad tasks:
 \begin{itemize}
-  \item filtering the dataset down to an amenable subset;
+  \setlength\itemsep{0em}
+  \item filtering the data set down to an amenable subset;
   \item understanding the types of function arguments and returns;
-  \item comparing types across different functions.
+  \item comparing type signatures across different functions.
 \end{itemize}
@@
 %to convey the relative scale of different packages.
-For exploring and filtering the dataset, \typevis relies on the treemap idiom
-to represent R packages and functions, with the block sizes conveying
-the relative popularity in the dataset.
-In addition to the treemaps, a search box can be used to directly filter
-by functions.
-Once a specific subset of one or more functions is selected,
-their type signatures are represented as flows over types
-where the width of the flow is proportional to frequency.
+For exploring and filtering the data set,
+\typevis relies on treemaps to represent R packages and functions,
+with node size conveying relative popularity.
+Additionaly, one can use a search box to directly filter
+by function name.
+Once a subset of one or more functions is selected,
+their type signatures are encoded as flows over types,
+where flow width is proportional to frequency.
 These flows can be readily compared within the same function,
-or across different functions.
+or between different functions.
 
-Tasks enabled by \typevis can assist language designers
-during multiple phases of development.
-For example, exploratory data analysis can identify unexpected uses
-of specific functions, or weed out language designs
-that are incompatible with existing programs.
-In particular, the visualization is built to answer questions
-relevant to the development of a gradual type system for R.
+\typevis can assist language designers during multiple phases of development.
+For example, exploratory data analysis can identify unexpected edge cases,
+or weed out language designs that are incompatible with existing programs.
+\typevis was built in support of a new gradual type system for R,
+whose adoption depends on integrating well with existing R code.
 However, the visualization is not limited to R and can be used to
-analyze type usage in any programming language where type information
-is available at run-time.
+analyze type usage in any language where similar data is available.
+
 Supplemental material, including the data and source code of
@@
 and type tags for each argument and return value.
+For example, for the function call \code{!(FALSE)},
+the recorded type signature would be \code{logical->logical},
+meaning that function \code{!} took a logical value and returned a logical value.
 The type tags are assigned by the dynamic analysis and
@@
 because R does not have native support for scalars.
-Our dataset consists of a table of reduced execution traces:
+Our data set consists of a table of reduced execution traces:
 because \typevis is designed for the analysis of types, not all information
@@
 that permits real-time exploration,
-even with large datasets.
-The default dataset contains over a million rows
+even with large data sets.
+The default data set contains over a million rows
 based on the analysis of almost five hundred
@@
 level deep.
-However, if the dataset was more deeply nested,
+However, if the data set was more deeply nested,
 \typevis can be configured to
@@
 of run-time type information for R.
-It supports real-time exploration of a dataset of
+It supports real-time exploration of a data set of
 execution traces from a large corpus of R packages.
@@
 and is suitable to use for any dynamically typed language
-with a similar dataset.
+with a similar data set.
 Additionally, the visualization could yield insights

### 2020-06-21 paper/main.tex
paper: My edits upto 4

@@
 with node size conveying relative popularity.
-Additionaly, one can use a search box to directly filter
-by function name.
+Additionally, one can search directly for a function by name.
 Once a subset of one or more functions is selected,
@@
 \typevis was built in support of a new gradual type system for R,
-whose adoption depends on integrating well with existing R code.
+whose adoption depends on integrating well with existing code.
 However, the visualization is not limited to R and can be used to
@@
 Supplemental material, including the data and source code of
-\typevis, is freely available: \todo{Fill with archival link}
+\typevis, is freely available: \todo{Add archival link}
 
@@
 
-R is a dynamically typed programming language widely used for
-statistical computing. Unlike statically typed languages,
-R programs do not contain type annotations
-and never pass through a typechecker.
+R is a lazy, multi-paradigm, dynamically-typed programming language
+widely used for statistical computing~\cite{morandat:2012}.
+Unlike statically-typed languages,
+R code contains no type annotations
+and never passes through a typechecker.
 At run-time,
@@
 
-Being a dynamically typed language most actively used in a single domain,
+Being a dynamically-typed language
+with widespread use in a single domain,
 R is a good target for a gradual type system.
-For statically typed languages, the type system is usually
+A gradual type system permits the integration of
+static and dynamic typing~\cite{siek:2006,tobin-hochstadt:2006}.
+Such a type system is capable of seamlessly mixing typed and untyped regions of code.
+
+For statically-typed languages, the type system is usually
 an inseparable part of the language. Since they are designed
-at the same time, the (static) type system precedes any code written
+at the same time, the static type system precedes any code written
 in the language.
-A gradual type system, by contrast,
-permits the integration of static and dynamic typing.
-Such a type system is capable of seamlessly mixing typed and untyped regions of code.
-Often, a gradual type system is designed atop
-an existing dynamically typed language.
-Since many R programs already exist,
-a successful gradual type system must be capable of
-typechecking idiomatic R code.
-\typevis aims to clarify how existing R programs
-use types to assist in this task.
-While this is our motivation, the visualization's use
-is not necessarily limited to this application.
-Anyone who has an interest in how the real R programs use types
-may find it useful.
+Gradual type systems, by contrast, are often designed atop
+an existing dynamically-typed language~\cite{tobin-hochstadt:2016}.
+Many R programs already exist,
+so a successful gradual type system must
+work well with extant code.
 
 We are not aware of any previous work that applies
-visualization techniques for the purposes of programming language design.
+visualization techniques for the purpose of programming language design.
 However, there are existing publications on understanding
@@
 Bohnet et al.~\cite{bohnet:2009} visualize massive program execution traces
-to assist developers in code comprehension of large software systems.
-Their tool applies pruning and summarization to reduce trace size.
-The visualization focuses on call topology,
+to improve developer comprehension of large software systems.
+Their tool applies pruning and summarization techniques to reduce trace size.
+They focus on call topology,
 describing the structure of function invocations,
@@
 rather than individual execution traces.
-Our visualization can be thought of as a hybrid of the two,
+\typevis has elements of both,
 investigating the system as a whole,
@@
 LaToza and Myers~\cite{latoza:2011} adorn nodes and edges with extra information,
-allowing the user gain further understanding of the system
+allowing the user to gain further understanding of the system
 under examination.
-This includes indicating methods that come from the same class---a
-basic form of type information.
+This includes grouping methods that come from the same class---a
+basic form of type data.
 
@@
 
-\todo{Alexi's paper says 412, but I see more packages in our data.}
-
-The data is provided by Alexi Turcotte:
-it consists of traces from 412 packages
+The data set we used comes from
+Turcotte et al.~\cite{turcotte:2020}
+and consists of traces from 412 packages,
 containing over $760,000$ lines of R code
@@
 the most popular R code repository.
-Packages hosted on CRAN must satisfy certain criteria
-to ensure they are high quality.
-Additionally, the subset of 412 packages chosen here
-all had code coverage above 65\%.\todo{I don't understand what subset? Should we remove this line?}
+Packages hosted on CRAN are curated and satisfy
+certain quality standards.
+Out of the over $15,000$ packages
+distributed on CRAN,
+the subset of 412 packages considered here
+all had code coverage above 65\%.
+\todo{Confirm with Alexi this paragraph. Get proper BibTeX entry.}
 
-Each execution trace was recorded by a dynamic analysis
-running the test, example, and vignette code of each package.
+Execution traces were recorded by running a dynamic analysis
+over the test, example, and vignette code of each package.
 A vignette is a form of documentation that weaves
 together prose and executable code.
-An execution trace is a list of ``raw'' function call signatures
+An execution trace is a list of function type signatures
 containing, among other attributes, the function name
 and type tags for each argument and return value.
-For example, for the function call \code{!(FALSE)},
-the recorded type signature would be \code{logical->logical},
-meaning that function \code{!} took a logical value and returned a logical value.
-The type tags are assigned by the dynamic analysis and
-refine over actual type tags in R; for example, the analysis
-assigns scalar tag \code{double} to a singleton vector of double values
+For example, the function call {\tt !(FALSE)}
+would be recorded as the type signature {\tt logical $\to$ logical},
+meaning that the function {\tt !} took a logical value
+and returned a logical value.\footnote{
+  In R, a logical value is a boolean.
+}
+The type tags are assigned by the dynamic analysis
+and refine the actual type tags in R.
+For instance, the analysis
+assigns the scalar tag {\tt double} to a singleton vector of double values
 because R does not have native support for scalars.
-Our data set consists of a table of reduced execution traces:
-because \typevis is designed for the analysis of types, not all information
-in the traces is relevant.
+The data set is a table of reduced execution traces.
+Because \typevis was designed for the analysis of types,
+not all information in the traces is relevant.
 %that have been processed to reduce some fields.
@@
 While \typevis has only been instantiated with
-data from R execution traces, the architecture is not
-limited to R.
-Any dynamically typed language %with the right run-time
+data from R execution traces, the architecture is not limited to R.
+Any dynamically-typed language, %with the right run-time
 %hooks could generate data for the tool.
-(or a statically typed language with run-time tag information)
+or statically-typed language with run-time tag information,
 could be instrumented to generate data for the tool.
-As long as execution trace information adheres to
-the same database schema, \typevis will
-work properly without any modifications.
+As long as the trace output adheres to
+the correct schema, \typevis will work properly without any modifications.
 
@@
 
-We interviewed the primary researcher developing R's gradual type system
+We interviewed the primary researcher developing a gradual type system for R
 to inventory relevant domain-specific tasks.
 Additionally, we sought feedback throughout the creation of \typevis
-to ensure it was adequately satisfying requirements.
-The domain-specific goals can be reduced to:
+to ensure it was adequately meeting requirements.
+The domain-specific goals identified were:
 
@@
 \item {\bf Find Function.} Quickly search for a particular function and identify basic information. This includes how often the function is called and which package contains it.
-\item {\bf Determine Types.} Given a specific function, determine the argument and return types of the function.
-Additionally, identify the frequency of a type signature.
-A user should be able to infer, for example, if the function is polymorphic or if it's a predicate.
-\item {\bf Compare Signatures.} Compare the occurrence of type signatures within a specific function, but also across many different ones.
-Users must be able to isolate individual signatures and see them in the context of other signatures and functions.
+\item {\bf Determine Types.} Given a specific function, determine the argument types, return types, and identify the frequency of specific type signatures.
+A user should be able to readily infer, for example, if a function is polymorphic or if it is a predicate.
+\item {\bf Compare Signatures.} Compare the occurrence of type signatures within a specific function, and across many different functions.
+Users must be able to isolate individual signatures and see them in the context of others.
 \end{enumerate}
@@
 See table~\ref{tab:tasks} for a categorization of these domain tasks
-the mid-level and low-level abstract task taxonomy
+with the mid-level and low-level abstract task taxonomy
 established by Brehmer and Munzner~\cite{brehmer:2013}.

### 2020-06-21 paper/main.tex
paper: More edits

@@
 
+\bgroup
+\def\arraystretch{1.75}
+\begin{table*}
+  \centering
+  \begin{tabularx}{\linewidth}{c|c|c|X}
+    Domain Goals & Search Task & Query Task & \multicolumn{1}{c}{Abstract Task Description} \\
+    \hline
+    Find Function & Locate & Identify & Finding a function is a location task, where the target function is known, but its location within the package hierarchy is not.
+Once the desired function is found, a user must be able to identify data of interest.
+\\
+
+    Determine Types & Browse & Identify & Determining types is a browse task, where the target type signature is unknown, but the location of the function of interest is known.
+While browsing, a user must be able to identify the frequencies of particular type signatures.\\
+
+    Compare Signatures & Lookup & Compare & Comparing type signatures is a lookup task where both target type signatures have already been located.
+A user must be able to discern how often a signature is used compared to another, both within a function and across different functions.\\
+  \end{tabularx}
+  \caption{Domain goals and their mid-level and low-level task abstractions.}
+  \label{tab:tasks}
+\end{table*}
+\egroup
+
 %%
@@
 
-\bgroup
-\def\arraystretch{1.75}
-\begin{table*}
-  \centering
-  \begin{tabularx}{\linewidth}{c|c|c|X}
-    Domain Goals & Search Task & Query Task & \multicolumn{1}{c}{Abstract Task Description} \\
-    \hline
-    Find Function & Locate & Identify & Finding a function is a location task, where the target function is known, but its location within the package hierarchy is not.
-Once the desired function is found, a user must be able to identify data of interest.
-\\
-
-    Determine Types & Browse & Identify & Determining types is a browse task, where the target type signature is unknown, but the location of the function of interest is known.
-While browsing, a user must be able to identify the frequencies of particular type signatures.\\
-
-    Compare Signatures & Lookup & Compare & Comparing type signatures is a lookup task where both target type signatures have already been located.
-A user must be able to discern how often a signature is used compared to another, both within a function and across different functions.\\
-  \end{tabularx}
-  \caption{Domain goals and their mid-level and low-level task abstractions.}
-  \label{tab:tasks}
-\end{table*}
-\egroup
-
-\section{Data and Collection}
+\section{Data and Collection} \label{sec:data}
 
@@
 would be recorded as the type signature {\tt logical $\to$ logical},
-meaning that the function {\tt !} took a logical value
-and returned a logical value.\footnote{
-  In R, a logical value is a boolean.
-}
+meaning that the function {\tt !} took a logical value (i.e. a boolean)
+and returned a logical value.
 The type tags are assigned by the dynamic analysis
@@
 \item {\bf Find Function.} Quickly search for a particular function and identify basic information. This includes how often the function is called and which package contains it.
-\item {\bf Determine Types.} Given a specific function, determine the argument types, return types, and identify the frequency of specific type signatures.
-A user should be able to readily infer, for example, if a function is polymorphic or if it is a predicate.
+\item {\bf Determine Types.} Given a specific function,
+  determine its argument and return types, and identify the frequency of specific type signatures.
+  A user should be able to readily infer, for example,
+  if a function is polymorphic (i.e. accepts arguments of multiple different types) or if it is a predicate (i.e. returns a logical value).
 \item {\bf Compare Signatures.} Compare the occurrence of type signatures within a specific function, and across many different functions.
-Users must be able to isolate individual signatures and see them in the context of others.
+A user must be able to isolate individual signatures and exmaine them in the context of others.
 \end{enumerate}
@@
 even with large data sets.
-The default data set contains over a million rows
-based on the analysis of almost five hundred
-different R packages.
-Figure~\ref{fig:typevis} displays
+The default data set,
+described in \S\ref{sec:data},
+contains over a million rows.
+Figure~\ref{fig:typevis} shows
 the heart of \typevis
@@
 Most aspects of the system can be tweaked
-by a user if desired.
+by the user if desired.
 
@@
 Selecting a package
-in the top treemap will immediately narrow the
-bottom view to only functions available in that package.
-Additionally, entering a function name in the search
+in the top treemap will cause the
+bottom view to show only functions available in that package.
+Entering a function name in the search bar
 will select it in the function treemap.
@@
 the search bar supports autocomplete
-that can quickly select a function among thousands.
+that can select one function among thousands.
 However, if one wants to browse
@@
 the treemap view is more applicable.
-Nodes in the bottom treemap convey the log-scaled
+Area in the bottom treemap conveys the log-scaled
 frequency of function calls,
@@
 elements of the treemap legible,
-even if absolute size differences are significant.
+even when absolute size differences are significant.
 
@@
 hierarchies.
-R packages are not structured deeply
-so the treemaps in our data are only one
-level deep.
-However, if the data set was more deeply nested,
-\typevis can be configured to
-handle that.
+The packages in our data are not deeply structured,
+so the treemaps in figure~\ref{fig:typevis} are only one level deep.
+However, if the data set recorded deeply nested packages,
+\typevis could be modified to handle that.
 
 If configured, \typevis allows for multiple function selection.
-Both the {\tt length.unit} and {\tt valid.unit} functions
-from the {\tt grid} package are selected in figure~\ref{fig:typevis}.
+In this example, the {\tt length.unit} and {\tt valid.unit} functions
+from the {\tt grid} package are selected.
 These functions are rendered simultaneously in the type flow
-panel, allowing one to make comparisons between them.
+panel, allowing the user to make comparisons between them.
 
@@
 
-Prominently featured as the center visualization,
-a parallel sets diagram encodes type signatures of the selected
+Prominently featured in the center of figure~\ref{fig:typevis}
+is a parallel sets diagram encoding the type signatures of the selected
 functions as flows.
@@
 are ideally suited for type signatures.
-Each data point is can be thought of as
+Each data point can be thought of as
 a tuple $(f, \tau_1, \ldots, \tau_n)$
@@
 and return type.
-Every component is categorical,
-and we are interested in visualizing frequencies of these points.
+Every component of that tuple is categorical,
+and we are interested in visualizing the frequencies of these points.
 This is the exact use-case for parallel sets.
 
-Flows begin at nodes labeling a function,
+Flows begin at a node labeling a function,
 curve across each argument type in order,
@@
 Flow segments are filled with a hue-varying color scale,
-based on the type where the segment ends.
-Beneath each node label is an approximate count.
-If the node represents a function, then the
-number is how many times the function was invoked.
-If the node represents a type, then the
-number is how often a value of that type
-at that argument position was recorded during analysis.
+determined by the type where the segment ends.
+Beneath each node label is an approximate count of either:
+how many times the function was invoked if the node represents a function,
+or how often a value of that type was recorded during analysis if the node represents a tuple.
 
-For example, in figure~\ref{fig:typevis},
+In figure~\ref{fig:typevis},
 the {\tt length.unit} function is called with
@@
 
-Interactivity significantly augments the
-usability of the parallel sets visualization.
+Interactivity significantly enhances the
+usability of the parallel sets diagram.
 Hovering over a flow will highlight the
 entire path of the flow
-and will fade other functions
-into the background.
+and will fade out all other functions.
 When displaying many flows at the same time,
-highlighting becomes especially critical
+highlighting becomes critical
 for user comprehension.
 Additional quantitative information
-about the flow is also supplied
-while hovering.
+about the flow is supplied on-demand while hovering.
 Clicking a flow will focus on that function
@@
 
-Flow layout is a critical consideration
-for a sufficiently legible visualization.
+Flow layout is a necessary consideration
+for a legible visualization~\cite{sugiyama:1981}.
 Figure~\ref{fig:decross} exemplifies what
-happens if flow layout is not adequately handled.
+happens if flow layout is not adequately computed.
 The top layout was generated without any
 flow decrossing algorithm,
-while the bottom layout does
-apply decrossing.
+while the bottom layout applies optimal decrossing.
 Even with highlighting,
-the utility of the visualization is vastly
-reduced without some decrossing effort.
+the visualization is almost impossible to understand
+due to excessive flow overlap.
 To maintain real-time performance,
@@
 First, the system attempts to construct
-a layout with the globally optimal
-minimum amount of flow crossing by
+a layout with the globally optimal solution,
+minimizing the amount of flow crossing by
 solving a mixed-integer linear program.
@@
 sized graphs this can take too long to compute.
-Therefore, after about $1$ second
+Therefore, after about $1$ second,
 the optimal algorithm is terminated
 if it has not finished, and a faster
-algorithm that only locally minimizes crossing
-between layers computes the layout.
+algorithm that only minimizes local crossing between layers
+computes the layout.
+While not ideal, it is much better than no decrossing
+at all.
 
@@
 
-Since the amount of data that a user may query is large,
+Since the amount of data a user may query is large,
 the application consists of both a frontend and a backend.
+This avoids excessive memory consumption and speeds up computations,
+as the client queries data from the server on an need-to-know basis.
 
 On the frontend, state management, interactivity,
-and DOM manipulation is all done within the Vue framework.
-This is in contrast with many web-based visualizations
+and DOM manipulation is all done with the Vue framework.
+This differs from many web-based visualizations
 that use D3 for DOM manipulation.
-We chose this setup because of the many different
-individual components whose state must be kept
-synchronized.
+We chose this setup because \typevis retains many distinct
+components whose state must be kept synchronized.
 This includes not only the visualizations, but also
-user interface elements like the autocomplete search bar.
+user interface elements like the search bar.
 In our experience, Vue's model of reactivity is
 more successful at maintaining consistency across the
-entire system than using D3 and vanilla JavaScript.
+entire application than D3 and vanilla JavaScript.
 While \typevis does not use D3 for DOM manipulation,
-it still does use D3 for other visualization-related
-computations.
+it still does use it for other visualization-related
+tasks. Decrossing is computed using web workers
+to allow early termination if the thread takes too long.
 
 Our backend is a Node.js application that serves
-JSON in a standard REST-style architecture.
+JSON with a standard REST-style architecture.
 Data is stored in a SQLite database that has been
 indexed to achieve lookups as fast as possible.
-To avoid excessive memory consumption and speed up computations,
-clients query data from the server on an as-needed basis.
-Thus, a stable network connection is required for
+A stable network connection is required for
 real-time interaction.
@@
 
-\typevis suffers from a number of shortcomings
-and does not fully address the entire range of possible tasks
+\typevis suffers from some shortcomings
+and does not fully address the entire range of tasks
 a language designer may need to perform.
 
-In particular, R supports a wide range of data types
-including types that contain dimensionality information.
+In particular, R supports a range of data types
+including ones containing dimensionality information.
 For example, an R array may have type {\tt integer[3]}
 indicating that it is an array of length $3$.
-This will get compressed down into a single array
-type labeled {\tt integer[]}.
+This will get compressed down into a single array of
+type {\tt integer[]}.
 Reducing the number of types can make the visualization
-more understandable---at the expensive of losing some
+more understandable, at the expensive of losing
 precision in type information.
 Ideally a user would be able to interactively tune
-the level of type granularity based on their interest
-in a specific property.
+the level of granularity based on their interest.
 Such a feature would be especially helpful for R
-since values can come attached with relevant additional
+since values can come attached with lots of relevant
 data such as the dynamic dispatch method
@@
 
-Filtering is an central mechanism in the system,
+Filtering is a central mechanism of our system,
 but it also comes at a cost.
@@
 in mind, then \typevis provides a useful
-local view of that type information.
-However, it does not support any kind of global
-view of type information across a large amount of
-functions. When the number of flows or nodes
+local view of that information.
+It does not support, however, any kind of global
+view of type information across a large number of
+functions. When the amount of flows or nodes
 becomes too great, the visualization will start
-paginating, making comparisons over large amounts
-of data impossible.
+paginating, making comparisons across all flows impossible.
 Aggregation and summarization would be key to
-making a global view of the data feasible---that
-remains future work.
+making a global view of the data feasible, but that remains future work.
 
@@
 We have presented \typevis, an interactive visualization
-of run-time type information for R.
-It supports real-time exploration of a data set of
-execution traces from a large corpus of R packages.
-To narrow the data under consideration,
-\typevis provides facilities for filtering
-and searching.
-Type signatures are represented as flows traversing across
+of run-time type information
+that supports real-time analysis and exploration.
+The tool was developed for a massive data set of
+execution traces from a corpus of popular R packages.
+
+\typevis provides facilities for filtering, searching,
+understanding, and comparing type information.
+Signatures are represented as flows traversing
 each argument type and return type in sequence.
-Flow widths are proportional to how often
-that type signature was seen during the dynamic analysis.
 Interactions provide additional details and
-assist users in comprehension and navigation.
+assist users with comprehension and navigation.
 
-Our tool is aimed at elucidating how R programmers
-use the language in practice.
+Our system is aimed at elucidating how R programmers
+use the language in the field.
 Specifically, the data and visualization will be used to
-inform the design of a gradual type system for R.
+inform the design of a future gradual type system for R.
 However, \typevis is not tied to the R ecosystem
-and is suitable to use for any dynamically typed language
-with a similar data set.
-Additionally, the visualization could yield insights
-useful for other purposes than gradual typing.
+and is suitable to use for any dynamically-typed language.
+The visualization could yield insights
+useful for purposes other than gradual typing,
+and beyond what we have considered.
 
 We hope that, in the future, programming language designers
-will collect and use empirical data about how their languages
-are used in practice.
-Such data can be used to make informed decisions about
-language design.
+will collect and use empirical data about their languages.
+Such data could be used to make informed decisions about design,
+where the cost of making a mistake is high.
 Visualization will be unavoidable in these scenarios
@@
 Generic, canned visualizations will not be sufficient.
-\typevis is just the first step in this direction.
+\typevis is a first step in this direction.
 
 \acknowledgments{
-  The authors wish to thank the students of CS 7250 for useful design feedback,
-  and contributors to all the open source projects that \typevis depends on.
+  The authors wish to thank
+  Alexi Turcotte,
+  Cody Dunne,
+  Sara Di Bartolomeo,
+  Younes El Idrissi Yazami,
+  Peta Maj,
+  the students of CS 7250,
+  and contributors to all the open source projects that \typevis uses.
   This work was supported in part by a Northeastern Graduate Fellowship.

### 2020-06-22 paper/main.tex
paper: More edits

@@
 
-\documentclass{vgtc}                          % final (conference style)
-%\documentclass[review]{vgtc}                 % review
+%\documentclass{vgtc}                          % final (conference style)
+\documentclass[review]{vgtc}                 % review
 %\documentclass[widereview]{vgtc}             % wide-spaced review
@@
 \usepackage{todonotes} %TODO: Remove after all TODO's are gone
+\usepackage{lineno}
+\linenumbers
 
@@
 \newcommand{\typevis}{{\sc TypeVis}\xspace}
+\newcommand{\figref}[1]{Figure~\ref{#1}\xspace}
+\newcommand{\pkgNumAnalyzed}{412\xspace}
+\newcommand{\settingsTab}{{\sc Settings}\xspace}
 
@@
 However, the visualization is not limited to R and can be used to
-analyze type usage in any language where similar data is available.
+analyze type usage in any language where similar data are available.
 
@@
 A gradual type system permits the integration of
-static and dynamic typing~\cite{siek:2006,tobin-hochstadt:2006}.
-Such a type system is capable of seamlessly mixing typed and untyped regions of code.
+static and dynamic typing~\cite{siek:2006,tobin-hochstadt:2006},
+and is capable of seamlessly mixing typed and untyped regions of code.
 
@@
 an inseparable part of the language. Since they are designed
-at the same time, the static type system precedes any code written
+at the same time, the type system precedes any code written
 in the language.
@@
 However, there are existing publications on understanding
-software systems through visualization.
+software through visualization.
 Bohnet et al.~\cite{bohnet:2009} visualize massive program execution traces
@@
 Turcotte et al.~\cite{turcotte:2020}
-and consists of traces from 412 packages,
+and consists of traces from \pkgNumAnalyzed packages,
 containing over $760,000$ lines of R code
@@
 The type tags are assigned by the dynamic analysis
-and refine the actual type tags in R.
+and refine the actual type tags at run-time.
 For instance, the analysis
@@
 to inventory relevant domain-specific tasks.
-Additionally, we sought feedback throughout the creation of \typevis
+Additionally, we sought his and other researchers' feedback
+throughout the creation of \typevis
 to ensure it was adequately meeting requirements.
@@
 \typevis is an interactive web-based visualization
-that permits real-time exploration,
+that permits real-time data exploration,
 even with large data sets.
-The default data set,
+The default data set of R execution traces,
 described in \S\ref{sec:data},
 contains over a million rows.
-Figure~\ref{fig:typevis} shows
-the heart of \typevis
-in its initial configuration.
-Most aspects of the system can be tweaked
-by the user if desired.
+
+\figref{fig:typevis} shows
+the heart of \typevis---the type flow panel on the left
+and the function filtering panel on the right.
+Type signatures shown on the left reflect
+the subset of the data selected on the right.
+By default, if no subset is selected, the type flow
+panel shows several of the most frequent function type signatures
+in the entire data set.
+%in its initial configuration.
+Various aspects of the system,
+such as the number of type signatures shown at once,
+can be tweaked by the user if desired;
+for this, a separate \settingsTab tab is provided.
 
@@
 
-The right-hand panel of figure~\ref{fig:typevis}
+The right-hand panel of \figref{fig:typevis}
 contains three elements: a search bar,
@@
 and a function treemap.
-All three are linked.
+All three components are linked.
 Selecting a package
-in the top treemap will cause the
-bottom view to show only functions available in that package.
+in the top treemap causes the bottom treemap
+to show only functions defined in that package;
+additionally, the type flow panel gets updated
+with the most frequent function type signatures in the package.
+Selecting a function in the bottom treemap
+makes the type flow panel show only signatures
+of that function.
 Entering a function name in the search bar
-will select it in the function treemap.
+updates the type flow panel immediately.
 
-These different components support different modes of search.
+These different components support at least two modes
+of search: quick function lookup and package exploration.
 If a user has a specific function in mind,
-the search bar supports autocomplete
-that can select one function among thousands.
-However, if one wants to browse
-all available functions and packages,
+they can type its name into the search bar and see its
+type signatures straight away;
+the search bar supports autocomplete that is
+able to select one function among thousands.
+If, on the other hand, one wants to browse
+all available packages and functions,
 the treemap view is more applicable.
-Area in the bottom treemap conveys the log-scaled
-frequency of function calls,
-and correspondingly the top treemap
-is scaled based on invocations
-of functions defined in that package.
+The area of a block in the function treemap (package treemap)
+conveys the log-scaled
+frequency of calls to that function
+(to all functions defined in that package).
+%Area in the bottom treemap conveys the log-scaled
+%frequency of function calls,
+%and correspondingly the top treemap
+%is scaled based on invocations
+%of functions defined in that package.
 Log-scaling and pagination keep
-elements of the treemap legible,
+elements of the treemaps legible,
 even when absolute size differences are significant.
 
-Treemaps are often used for displaying hierarchical data~\cite{shneiderman:1992}.
-Package systems in many languages,
-like Java,
-contain deeply nested namespace hierarchies.
-This idiom is appropriate for navigation of such
-hierarchies.
+The data set can additionally be filtered by analyzed package,
+in other words, the package containing the call site of a function, not the definition site.
+As mentioned in \S\ref{sec:data}, the data set contains information
+from \pkgNumAnalyzed analyzed packages, which are all taken into account
+by default. However, a user can limit the set of analyzed packages,
+using a search bar in \settingsTab.
+
+Although treemaps are often used for displaying
+hierarchical data~\cite{shneiderman:1992},
+in \typevis they have a flat structure.
 The packages in our data are not deeply structured,
-so the treemaps in figure~\ref{fig:typevis} are only one level deep.
-However, if the data set recorded deeply nested packages,
+so the treemaps in \figref{fig:typevis} are only one level deep.
+However, in many programming languages, like Java,
+package systems contain deeply nested namespace hierarchies.
+The treemap idiom is appropriate for navigation of such
+hierarchies, and
+%However, if the data set recorded deeply nested packages,
 \typevis could be modified to handle that.
 
-If configured, \typevis allows for multiple function selection.
-In this example, the {\tt length.unit} and {\tt valid.unit} functions
-from the {\tt grid} package are selected.
+By default, \typevis allows only one function to be selected at a time,
+but it can be configured for multiple function selection.
+In the example from \figref{fig:typevis},
+the {\tt length.unit} and {\tt valid.unit} functions
+from the {\tt grid} package are both selected.
 These functions are rendered simultaneously in the type flow
 panel, allowing the user to make comparisons between them.
+However, if the user wants to explore different functions rather than
+compare them, the default one-function mode is more appropriate.
 
@@
 
-Prominently featured in the center of figure~\ref{fig:typevis}
+Prominently featured in the center of \figref{fig:typevis}
 is a parallel sets diagram encoding the type signatures of the selected
@@
 Each data point can be thought of as
-a tuple $(f, \tau_1, \ldots, \tau_n)$
+a tuple $(f, \tau_1, \ldots, \tau_n, \tau_r)$
 corresponding to the function name,
@@
 how many times the function was invoked if the node represents a function,
-or how often a value of that type was recorded during analysis if the node represents a tuple.
+or how often a value of that type was recorded during analysis if the node represents a type.
 
-In figure~\ref{fig:typevis},
+In \figref{fig:typevis},
 the {\tt length.unit} function is called with
@@
 entire path of the flow
-and will fade out all other functions.
+and will fade out all other functions,
+as shown in \figref{fig:typevis}.
 When displaying many flows at the same time,
@@
 Clicking a flow will focus on that function
-only, filtering away all others.
+only, filtering away all others;
+double clicking outside flows will bring back the previous view.
+The number of flows shown at a time is limited,
+and will paginate if that threshold is exceeded.
 
@@
 Even with highlighting,
-the visualization is almost impossible to understand
+the top visualization is almost impossible to understand
 due to excessive flow overlap.
@@
 minimizing the amount of flow crossing by
-solving a mixed-integer linear program.
+solving a mixed-integer linear program~\cite{ebner:2005}.
 Unfortunately, for even moderately
@@
 the optimal algorithm is terminated
-if it has not finished, and a faster
-algorithm that only minimizes local crossing between layers
-computes the layout.
-While not ideal, it is much better than no decrossing
-at all.
+if it has not finished, and the final layout is computed
+by a faster
+algorithm that only minimizes local crossing between layers.
+%computes the layout.
+While the resulting layout might not be ideal,
+it is much better than the one without decrossing at all.
 
@@
 Since the amount of data a user may query is large,
-the application consists of both a frontend and a backend.
+the application is split into a frontend and a backend.
+The backend is responsible for storing the data
+and fetching required information from the underlying database.
+The frontend serves as an interactive interface to the data
+and provides the visualization.
 This avoids excessive memory consumption and speeds up computations,
-as the client queries data from the server on an need-to-know basis.
+as the client queries data from the server on a need-to-know basis.
 
@@
 
+Designed with a specific use case in mind,
 \typevis suffers from some shortcomings
@@
 
-In particular, R supports a range of data types
+In particular, \typevis does not allow for exploring 
+type information at differing granularities.
+R supports a range of data types
 including ones containing dimensionality information.
@@
 indicating that it is an array of length $3$.
-This will get compressed down into a single array of
+Every such type will get compressed down %into a single array of
+to a simpler array
 type {\tt integer[]}.
@@
 precision in type information.
-Ideally a user would be able to interactively tune
+Ideally, a user would be able to interactively tune
 the level of granularity based on their interest.
-Such a feature would be especially helpful for R
-since values can come attached with lots of relevant
+Such a feature would be especially helpful for R,
+because values can come attached with lots of relevant
 data such as the dynamic dispatch method
@@
 that supports real-time analysis and exploration.
-The tool was developed for a massive data set of
+The tool was instantiated with a massive data set of
 execution traces from a corpus of popular R packages.
@@
 understanding, and comparing type information.
-Signatures are represented as flows traversing
+Type signatures of function calls
+are represented as flows traversing
 each argument type and return type in sequence.
-Interactions provide additional details and
+Quick function lookup is supported by a search bar,
+with package and function treemaps facilitating
+function exploration.
+Various interactions provide additional details and
 assist users with comprehension and navigation.
@@
 However, \typevis is not tied to the R ecosystem
-and is suitable to use for any dynamically-typed language.
+and is suitable to use for any %dynamically-typed language.
+language that supports run-time type information.
 The visualization could yield insights
@@
 We hope that, in the future, programming language designers
-will collect and use empirical data about their languages.
+will collect and use empirical data about their languages
+more extensively.
 Such data could be used to make informed decisions about design,

### 2020-07-07 src/server/api.js
add support for the secod database

@@ -8,9 +8,13 @@ let router = express.Router();
 //
 // Database
 //
-const url = process.env.TYPEVIS_DB;
+const url    = process.env.TYPEVIS_DB;       // base information
+const urlDet = process.env.TYPEVIS_DB_DET ?  // detailed information
+  process.env.TYPEVIS_DB_DET : 
+  process.env.TYPEVIS_DB;
 const sqlite3 = require("sqlite3").verbose();
-let db = new sqlite3.Database(url);
+let db    = new sqlite3.Database(url);
+let dbDet = new sqlite3.Database(urlDet);
 
 //
 // Queries
@@ -47,8 +51,9 @@ const FALSE = "0 = 1";
 const OR = (clause, n, empty) =>
   n === 0 ? empty : "(" + FALSE + ` OR ${clause}`.repeat(n) + ")";
 
-function query(sql, params, f) {
-  db.all(sql, params, (err, rows) => {
+function query(getDetails, sql, params, f) {
+  const dbToUse = getDetails ? dbDet : db;
+  dbToUse.all(sql, params, (err, rows) => {
     if (err) {
       throw err;
     }
@@ -67,14 +72,14 @@ module.exports = (app, server) => {
   // GET /api/analyzed
   //
   router.get("/api/analyzed", function(req, res) {
-    query(ANALYZED, [], names => res.json(names));
+    query(false, ANALYZED, [], names => res.json(names));
   });
 
   //
   // GET /api/analyzed
   //
   router.get("/api/stats", function(req, res) {
-    query(DATASET_STATS, [], items => res.json(items));
+    query(false, DATASET_STATS, [], items => res.json(items));
   });
 
   //
@@ -83,7 +88,7 @@ module.exports = (app, server) => {
   router.get("/api/all_funs", function(req, res) {
     const analyzed = req.query.analyzed || [];
     const where = OR(ANALYZED_EQ, analyzed.length, TRUE);
-    query(ALL_FUNS(where), analyzed, names => res.json(names));
+    query(false, ALL_FUNS(where), analyzed, names => res.json(names));
   });
 
   //
@@ -92,7 +97,7 @@ module.exports = (app, server) => {
   router.get("/api/pkgs", function(req, res) {
     const analyzed = req.query.analyzed || [];
     const where = OR(ANALYZED_EQ, analyzed.length, TRUE);
-    query(PKGS(where), analyzed, function(items) {
+    query(false, PKGS(where), analyzed, function(items) {
[…]

### 2020-07-07 src/store/index.js

@@ -95,6 +95,10 @@ const funs = [];
 // An array of results from querying function type information.
 const types = [];
 
+// Boolean
+// Flag indicating whether to request detailed data for type flows
+const details = false;
+
 //
 // Store
 //
@@ -115,7 +119,8 @@ export default new Vuex.Store({
     allFuns,
     funs,
     pkgs,
-    types
+    types,
+    details
   },
 
   mutations: {
@@ -169,6 +174,9 @@ export default new Vuex.Store({
     },
     types(state, val) {
       state.types = val;
+    },
+    details(state, val) {
+      state.details = val;
     }
   },
 
@@ -236,7 +244,8 @@ export default new Vuex.Store({
             analyzed: state.analyzed,
             pkg: state.selectedPkg,
             funs,
-            limit: state.defaultLimit
+            limit: state.defaultLimit,
+            details: state.details
           }
         })
         .then(response => {

### 2020-07-07 src/server/api.js
second database | server fix

@@ -120,6 +120,8 @@ module.exports = (app, server) => {
   // GET /api/types
   //
   router.get("/api/types", function(req, res) {
+    const details = req.query.details || false;
+
     const analyzed = req.query.analyzed || [];
     const limitVal = req.query.limit || DEFAULT_LIMIT;
     const whereAnalyzed = OR(ANALYZED_EQ, analyzed.length, TRUE);
@@ -136,7 +138,7 @@ module.exports = (app, server) => {
     const params = analyzed.concat(funs.flat()).concat(pkg);
     const QUERY = analyzed.length == 0 ? TYPES_ALL : TYPES;
 
-    query(req.query.details, QUERY(where) + limit, params,
+    query(details, QUERY(where) + limit, params,
       results => res.json(results)
     );
   });

### 2020-07-07 src/store/index.js

@@ -283,7 +283,13 @@ export default new Vuex.Store({
 
     setClearSelectedFunsOnPkg({ commit }, value) {
       commit("clearSelectedFunsOnPkg", value);
-    }
+    },
+
+    // Sets the details flag
+    setDetails({ commit, dispatch }, details) {
+      commit("details", details);
+      dispatch("queryTypes");
+    },
   },
 
   modules: {}

### 2020-07-07 src/server/api.js
second database | refactor server

@@ -42,6 +42,8 @@ const PKG_FUN_EQ = "(package = ? AND fun_name = ?)";
 const LIMIT = where => ` LIMIT ${where}`;
 const DEFAULT_LIMIT = 36;
 
+const DEFAULT_DETAILS = false;
+
 //
 // Util
 //
@@ -72,14 +74,14 @@ module.exports = (app, server) => {
   // GET /api/analyzed
   //
   router.get("/api/analyzed", function(req, res) {
-    query(false, ANALYZED, [], names => res.json(names));
+    query(DEFAULT_DETAILS, ANALYZED, [], names => res.json(names));
   });
 
   //
   // GET /api/analyzed
   //
   router.get("/api/stats", function(req, res) {
-    query(false, DATASET_STATS, [], items => res.json(items));
+    query(DEFAULT_DETAILS, DATASET_STATS, [], items => res.json(items));
   });
 
   //
@@ -88,7 +90,9 @@ module.exports = (app, server) => {
   router.get("/api/all_funs", function(req, res) {
     const analyzed = req.query.analyzed || [];
     const where = OR(ANALYZED_EQ, analyzed.length, TRUE);
-    query(false, ALL_FUNS(where), analyzed, names => res.json(names));
+    query(DEFAULT_DETAILS, ALL_FUNS(where), analyzed,
+      names => res.json(names)
+     );
   });
 
   //
@@ -109,7 +113,7 @@ module.exports = (app, server) => {
     const analyzed = req.query.analyzed || [];
     const pkg = req.query.pkg;
     const where = OR(ANALYZED_EQ, analyzed.length, TRUE) + " AND " + PKG_EQ;
-    query(false, FUNS(where), analyzed.concat([pkg]),
+    query(DEFAULT_DETAILS, FUNS(where), analyzed.concat([pkg]),
       function(items) {
         res.json(items);
       }
@@ -120,8 +124,8 @@ module.exports = (app, server) => {
   // GET /api/types
   //
   router.get("/api/types", function(req, res) {
-    const details = req.query.details || false;
-
+    const details = (req.query.details == 'true') || false;
+    
     const analyzed = req.query.analyzed || [];
     const limitVal = req.query.limit || DEFAULT_LIMIT;
     const whereAnalyzed = OR(ANALYZED_EQ, analyzed.length, TRUE);

### 2020-07-07 src/server/api.js
second database helper

@@ -38,6 +38,7 @@ const DATASET_STATS = `SELECT *  FROM stats`;
 const ANALYZED_EQ = "package_being_analyzed = ?";
 const PKG_EQ = "package = ?";
 const PKG_FUN_EQ = "(package = ? AND fun_name = ?)";
+const ARG_LIMITED = "(arg_t5 = 'NA')";
 
 const LIMIT = where => ` LIMIT ${where}`;
 const DEFAULT_LIMIT = 36;
@@ -74,7 +75,8 @@ module.exports = (app, server) => {
   // GET /api/analyzed
   //
   router.get("/api/analyzed", function(req, res) {
-    query(DEFAULT_DETAILS, ANALYZED, [], names => res.json(names));
+    const details = (req.query.details == 'true') || false;
+    query(details, ANALYZED, [], names => res.json(names));
   });
 
   //
@@ -138,7 +140,11 @@ module.exports = (app, server) => {
     const wherePkg = defaultData ? OR(PKG_EQ, pkg.length, TRUE) : TRUE;
 
     const limit = defaultData ? LIMIT(limitVal) : "";
-    const where = [whereAnalyzed, whereFuns, wherePkg].join(" AND ");
+    const whereClauses = [whereAnalyzed, whereFuns, wherePkg];
+    // to make preview nicer, limit signatures to certain number of arguments
+    if (defaultData)
+      whereClauses.push(ARG_LIMITED);
+    const where = whereClauses.join(" AND ");
     const params = analyzed.concat(funs.flat()).concat(pkg);
     const QUERY = analyzed.length == 0 ? TYPES_ALL : TYPES;
 

### 2020-07-07 src/store/index.js

@@ -5,6 +5,7 @@
 import Vue from "vue";
 import Vuex from "vuex";
 import axios from "axios";
+import lodash from "lodash";
 Vue.use(Vuex);
 
 //
@@ -249,8 +250,21 @@ export default new Vuex.Store({
           }
         })
         .then(response => {
+          // grou by function names before sorting
+          const groupedData = lodash.groupBy(response.data, "fun_name");
+          const groupedWeighedData = Object.keys(groupedData).map(function(k) {
+            return {
+              name: k,
+              data: groupedData[k],
+              count: groupedData[k].reduce((acc, d) => acc + d.count, 0)
+            };
+          });
+          let sortedData = groupedWeighedData
+            .sort((a, b) => b.count - a.count)
+            .map(a => a.data)
+            .flat();
           commit("decrementPending");
-          commit("types", response.data);
+          commit("types", sortedData);
         });
     },
 
@@ -289,7 +303,7 @@ export default new Vuex.Store({
     setDetails({ commit, dispatch }, details) {
       commit("details", details);
       dispatch("queryTypes");
-    },
+    }
   },
 
   modules: {}

### 2020-07-08 src/server/api.js
buggy

@@ -75,15 +75,15 @@ module.exports = (app, server) => {
   // GET /api/analyzed
   //
   router.get("/api/analyzed", function(req, res) {
-    const details = (req.query.details == 'true') || false;
-    query(details, ANALYZED, [], names => res.json(names));
+    query(false, ANALYZED, [], names => res.json(names));
   });
 
   //
-  // GET /api/analyzed
+  // GET /api/stats
   //
   router.get("/api/stats", function(req, res) {
-    query(DEFAULT_DETAILS, DATASET_STATS, [], items => res.json(items));
+    const details = (req.query.details == 'true') || false;
+    query(details, DATASET_STATS, [], items => res.json(items));
   });
 
   //

### 2020-07-08 src/store/index.js
Add flow justification and move detailed data to settings

@@ -40,6 +40,10 @@ const defaultLimit = 36;
 // Number of flows per page
 const flowsPerPage = 12;
 
+// Boolean
+// If the flow layout is justified or not.
+const flowsJustified = false;
+
 //
 // Boolean
 // If the Sankey diagram should be horizontal.
@@ -111,6 +115,7 @@ export default new Vuex.Store({
     autocompleteWithFuns,
     defaultLimit,
     flowsPerPage,
+    flowsJustified,
     horizontalLayout,
     clearSelectedFunsOnPkg,
     selectedPkg,
@@ -145,6 +150,9 @@ export default new Vuex.Store({
     },
     flowsPerPage(state, val) {
       state.flowsPerPage = val;
+    },
+    flowsJustified(state, val) {
+      state.flowsJustified = val;
     },
     horizontalLayout(state, val) {
       state.horizontalLayout = val;

### 2020-07-08 paper/main.tex
Update paper

@@
 
-%\documentclass{vgtc}                          % final (conference style)
-\documentclass[review]{vgtc}                 % review
+\documentclass{vgtc}                          % final (conference style)
+%\documentclass[review]{vgtc}                 % review
 %\documentclass[widereview]{vgtc}             % wide-spaced review
@@
 
+\usepackage[T1]{fontenc}% to use more chars & support copy & paste of special chars from PDF
+
 %% it is recomended to use ``\autoref{sec:bla}'' instead of ``Fig.~\ref{sec:bla}''
@@
 
+\usepackage{hyperref}
+\usepackage[noabbrev]{cleveref}
 \usepackage{caption}
@@
 \usepackage{todonotes} %TODO: Remove after all TODO's are gone
-\usepackage{lineno}
-\linenumbers
+
+%% ORCID
+\usepackage{scalerel}
+\usepackage{tikz}
+\usetikzlibrary{svg.path}
+
+\definecolor{orcidlogocol}{HTML}{A6CE39}
+\tikzset{
+  orcidlogo/.pic={
+    \fill[orcidlogocol] svg{M256,128c0,70.7-57.3,128-128,128C57.3,256,0,198.7,0,128C0,57.3,57.3,0,128,0C198.7,0,256,57.3,256,128z};
+    \fill[white] svg{M86.3,186.2H70.9V79.1h15.4v48.4V186.2z}
+                 svg{M108.9,79.1h41.6c39.6,0,57,28.3,57,53.6c0,27.5-21.5,53.6-56.8,53.6h-41.8V79.1z M124.3,172.4h24.5c34.9,0,42.9-26.5,42.9-39.7c0-21.5-13.7-39.7-43.7-39.7h-23.7V172.4z}
+                 svg{M88.7,56.8c0,5.5-4.5,10.1-10.1,10.1c-5.6,0-10.1-4.6-10.1-10.1c0-5.6,4.5-10.1,10.1-10.1C84.2,46.7,88.7,51.3,88.7,56.8z};
+  }
+}
+
+\newcommand\orcid[2]{\href{https://orcid.org/#2}{#1 \mbox{\scalerel*{
+\begin{tikzpicture}[yscale=-1,transform shape]
+\pic{orcidlogo};
+\end{tikzpicture}
+}{|}}}}
 
@@
 \newcommand{\code}[1]{\texttt{\small{#1}}}
-\newcommand{\typevis}{{\sc TypeVis}\xspace}
-\newcommand{\figref}[1]{Figure~\ref{#1}\xspace}
+\newcommand{\ourmethodplaintext}{\textsc{TyVis}\xspace}
+\newcommand{\ourmethod}{{\sc \ourmethodplaintext}\xspace}
 \newcommand{\pkgNumAnalyzed}{412\xspace}
 \newcommand{\settingsTab}{{\sc Settings}\xspace}
+
+\newcommand{\osfsupplementplain}{osf.io/mc6zt}
+\newcommand{\osfsupplement}{\formattedurl{https://\osfsupplementplain}{\osfsupplementplain}}
+\newcommand{\osfsupplementlong}{\formattedurl{https://\osfsupplementplain}{https://\osfsupplementplain}}
+
+\newcommand{\osfprereg}{\formattedurl{https://\osfpreregplain}{\osfpreregplain}}
+
+\newcommand{\formattedurl}[2]{\href{#1}{\small\texttt{#2}}\xspace}
 
@@
 
-\title{Type Visualization for R}
+\title{\ourmethodplaintext: Visualizing Function Type Signatures in R}
 
-\author{Cameron Moy\thanks{camoy@ccs.neu.edu}
-\and Julia Belyakova\thanks{belyakovay@ccs.neu.edu}}
+\newcommand{\mailtohref}[1]{\href{mailto:camoy@ccs.neu.edu;belyakovay@ccs.neu.edu;alexi@ccs.neu.edu;dibartolomeo.s@northeastern.edu;c.dunne@northeastern.edu}{\color{black}#1}}
+
+\author{
+    \orcid{Cameron Moy}{0000-0002-4384-6351}%
+    \thanks{%
+        E-mails: \mailtohref{%
+        [~%
+        {camoy} |
+        {belyakovay} |
+        {alexi}~]@ccs.neu.edu,
+        [~%
+        {dibartolomeo.s} |
+        {c.dunne}~]@northeastern.edu}
+        \vspace{0.55in}
+    }\\%
+    \and \orcid{Julia Belyakova}{0000-0002-7490-8500}\\%
+    \and \orcid{Alexi Turcotte}{0000-0002-0381-0477}\\%
+    \and \orcid{Sara Di Bartolomeo}{0000-0001-9517-3526}\\%
+    \and \orcid{Cody Dunne}{0000-0002-1609-9776}\\%
+}
 \affiliation{\scriptsize Northeastern University}
+
+\teaser{
+  \includegraphics[width=\textwidth]{img/teaser.png}
+  \caption{Type signatures for a subset of R's base package functions shown in the type flow visualization of \ourmethod.
+  The function name is listed at the top followed by the first two argument types.
+  All argument and return types are shown in the full visualization.}
+}
 
@@
   Data-driven approaches to programming language design are uncommon.
-  Despite the availability of large code repositories, distilling useful
-  information from programs remains difficult. Important dimensions,
-  like run-time type data, are inscrutable without the appropriate tools.
-  \typevis is an interactive visualization for exploring and analyzing
-  type information from program execution traces---aiding
-  user understanding of function type signatures across many executions.
-  The system uses flows to intuitively represent categorical
-  and numerical aspects of type signatures.
-  Insights derived from our visualization are aimed at informing
-  language design decisions---specifically of a new gradual type system
-  currently being developed for the R programming language.
+  Despite the availability of large code repositories, distilling semantically-rich information from programs remains difficult.
+  Important dimensions, like run-time type data, are inscrutable without the appropriate tools.
+  We contribute \ourmethod, an interactive visualization for exploring and analyzing type information from program execution traces --- aiding user understanding of function type signatures across many executions.
+  The system uses flows to intuitively represent categorical and numerical aspects of type signatures.
+  Insights derived from our visualization are aimed at informing language design decisions --- specifically of a new gradual type system currently being developed for the R programming language.
+  A full version of this paper with all supplemental material is available at \osfsupplementlong
 }
@@
 
-Programming languages commonly evolve by decree. Often, the language designer
-decides that a new feature is necessary, or that a past feature was
-ill-conceived. Thus, the language moves forward---
-forcing its users to adapt to the changes.
-Sometimes, these changes are the result of rigorous theory
-or the product of community feedback. However, rarely is language design
-informed by empirical data on how programs are \emph{actually
-written} in practice. Why so?
+\bgroup
+\def\arraystretch{1.75}
+\begin{table*}
+  \centering
+  \begin{tabularx}{\linewidth}{c|c|c|X}
+    \textbf{Domain Goals} & \textbf{Search Task} & \textbf{Query Task} & \multicolumn{1}{c}{\textbf{Abstract Task Description}} \\
+    \hline
+    Find function      & Locate & Identify & 
+    Finding a function is a locate task, where the target function is known, but its location within the package hierarchy is not.
+    Once the desired function is found, a user must be able to identify relevant data of interest. \\
 
-%For a popular language, data collection is no challenge. Vast
-Thanks to the prevalence of open source code, collecting data on
-the usage of a popular programming language is feasible. Vast
-quantities of publicly available code are hosted on repositories such as Github.
-Furthermore, if there is concern about the quality of this data,
-such as code duplication~\cite{lopes:2017},
-code can be obtained from curated
-language-specific package repositories like npm or CRAN.
+    Determine types    & Browse & Identify & 
+    Determining types is a browse task, where the target type signature is unknown, but the location of the function of interest is known.
+    While browsing, a user must be able to identify the frequencies of particular type signatures.\\
 
-To inform programming language design,
-the collected data needs to be analyzed and interpreted.
-Programs are complex and highly structured,
-so simple semantically-ignorant metrics like ``lines of code'' carry little significance.
-Therefore, researchers often employ static and dynamic analyses
-to gather information about specific aspects of programs.
-Even then, it may be difficult to make sense of the
-results of these analyses,
-especially if the data set is large.
-%Rather, the difficulty lies with analyzing and interpreting the data
-%at hand. Code is complex and highly structured. Simple
+    Compare signatures & Lookup & Compare  & 
+    Comparing type signatures is a lookup task where both target type signatures have already been located.
+    A user must be able to discern how often a signature is used compared to another, both within a function and across different functions.\\
+  \end{tabularx}
+  \caption{
+  Domain goals for programming language designers trying to understand function type signatures at run time.
+  We captured these goals from interviews and created mid-level and low-level task abstractions based on Brehmer \& Munzner's multi-level task typology~\cite{brehmer:2013}.}
+  \label{tab:tasks}
+\end{table*}
+\egroup
 
-To analyze just one aspect of programming languages, \emph{type usage}, we
-built \typevis, an interactive visualization of run-time type signatures.
-%an interactive visualization of run-time types for the language R.
-\typevis was created with the R programming language in mind,
-and builds on a data set of execution traces
-mined from the unit tests and vignettes
-of the most widely used libraries in the R ecosystem.
+Programming languages commonly evolve by decree.
+Often, the language designer decides that a new feature is necessary, or that a past feature was ill-conceived.
+Thus, the language moves forward --- forcing its users to adapt to the changes.
+Sometimes, these changes are the result of rigorous theory or the product of community feedback.
+However, rarely is language design informed by empirical data on how programs are \emph{actually written} in practice~\cite{hanenberg:2010}.
+Why so?
+
+Thanks to the prevalence of open source code, collecting data on the usage of a popular programming language is feasible.
+Vast quantities of code are publicly available online on language-specific package servers like npm for JavaScript or CRAN for R.
+Some repositories, like CRAN, only include high-quality libraries.
+Using a curated source of code mitigates some data collection concerns like code duplication~\cite{lopes:2017}.
+
+To inform programming language design, the collected data needs to be analyzed and interpreted.
+Programs are complex and highly structured, so simple semantically-unaware metrics like ``lines of code'' carry little significance.
+Therefore, researchers often employ static and dynamic analyses to gather information about specific aspects of programs.
+Even then, it may be difficult to make sense of the results of these analyses, especially if the data set is large.
+
+Following the design study ``lite'' methodology~\cite{syeda:2020}, we built \ourmethod, an interactive visualization of run-time type signatures.
+A type signature is a description of the argument and return types that a particular function is called with at run-time.
+\ourmethod was created with the R programming language in mind, and builds on a data set of execution traces mined from the unit tests and vignettes of the most widely used libraries in the R ecosystem.
 
@@
 \end{itemize}
-%For filtering, \typevis uses the treemap idiom
-%to convey the relative scale of different packages.
-For exploring and filtering the data set,
-\typevis relies on treemaps to represent R packages and functions,
-with node size conveying relative popularity.
+
+For exploring and filtering the data set, \ourmethod relies on treemaps to represent R packages and functions, with node size conveying relative popularity.
 Additionally, one can search directly for a function by name.
-Once a subset of one or more functions is selected,
-their type signatures are encoded as flows over types,
-where flow width is proportional to frequency.
-These flows can be readily compared within the same function,
-or between different functions.
+Once a subset of one or more functions is selected, their type signatures are encoded as flows over types, where flow width is proportional to frequency.
+These flows can be readily compared within the same function, or between different functions.
 
-\typevis can assist language designers during multiple phases of development.
-For example, exploratory data analysis can identify unexpected edge cases,
-or weed out language designs that are incompatible with existing programs.
-\typevis was built in support of a new gradual type system for R,
-whose adoption depends on integrating well with existing code.
-However, the visualization is not limited to R and can be used to
-analyze type usage in any language where similar data are available.
-
-Supplemental material, including the data and source code of
-\typevis, is freely available: \todo{Add archival link}
-
-\bgroup
-\def\arraystretch{1.75}
-\begin{table*}
-  \centering
-  \begin{tabularx}{\linewidth}{c|c|c|X}
-    Domain Goals & Search Task & Query Task & \multicolumn{1}{c}{Abstract Task Description} \\
-    \hline
-    Find Function & Locate & Identify & Finding a function is a location task, where the target function is known, but its location within the package hierarchy is not.
-Once the desired function is found, a user must be able to identify data of interest.
-\\
-
-    Determine Types & Browse & Identify & Determining types is a browse task, where the target type signature is unknown, but the location of the function of interest is known.
-While browsing, a user must be able to identify the frequencies of particular type signatures.\\
-
-    Compare Signatures & Lookup & Compare & Comparing type signatures is a lookup task where both target type signatures have already been located.
-A user must be able to discern how often a signature is used compared to another, both within a function and across different functions.\\
-  \end{tabularx}
-  \caption{Domain goals and their mid-level and low-level task abstractions.}
-  \label{tab:tasks}
-\end{table*}
-\egroup
+\ourmethod can assist language designers during multiple phases of development.
+For example, exploratory data analysis can identify unexpected edge cases, or weed out language designs that are incompatible with existing programs.
+\ourmethod was built in support of a new gradual type system for R, whose adoption depends on integrating well with existing code.
+A type system provides a conservative approximation of run-time types --- thus, insight into existing type usage can inform decision-making.
+Without a data-driven tool, gradual type system designers are left to guess about how their language is used in practice.
+Our aim is to eliminate such guesswork.
+While our visualization was created for a specific use case, it is not limited to R and can be used to analyze type usage in any language where similar data are available.
+Even beyond this domain, linking parallel sets and treemaps can be a powerful visualization technique.
+For example, parallel sets can display survey responses with many categorical questions~\cite{kosara:2006}, but a large number of responses can introduce too much visual clutter.
+Hierarchical filtering, with a treemap, can eliminate irrelevant information and recover an interpretable representation.
 
@@
 
-R is a lazy, multi-paradigm, dynamically-typed programming language
-widely used for statistical computing~\cite{morandat:2012}.
-Unlike statically-typed languages,
-R code contains no type annotations
-and never passes through a typechecker.
-At run-time,
-functions may be called with values of arbitrary type,
-and may return values of arbitrary type.\footnote{
+R is a lazy, multi-paradigm, dynamically-typed programming language widely used for statistical computing~\cite{morandat:2012}.
+Unlike statically-typed languages, R code contains no type annotations and never passes through a typechecker.
+At run-time, functions may be called with values of arbitrary type, and may return values of arbitrary type.\footnote{
   We use the word ``type'' to mean what others may call a ``type tag.''
-  This is a run-time concept---we do not assume a static type system
-  because there is none. Indeed, the goal is to help design one!
-  A ``type signature'' in this context is one particular configuration
-  of argument and return types a function is called with at run-time.
+  This is a run-time concept --- we do not assume a static type system   because there is none.
+  Indeed, the goal is to help design one!
+  A ``type signature'' in this context is one particular configuration of argument and return types a function is called with at run-time.
 }
 
-Being a dynamically-typed language
-with widespread use in a single domain,
-R is a good target for a gradual type system.
-A gradual type system permits the integration of
-static and dynamic typing~\cite{siek:2006,tobin-hochstadt:2006},
-and is capable of seamlessly mixing typed and untyped regions of code.
+Being a dynamically-typed language with widespread use in a single domain, R is a good target for a gradual type system.
+A gradual type system permits the integration of static and dynamic typing~\cite{siek:2006,tobin-hochstadt:2006}, and is capable of seamlessly mixing typed and untyped regions of code.
 
-For statically-typed languages, the type system is usually
-an inseparable part of the language. Since they are designed
-at the same time, the type system precedes any code written
-in the language.
-Gradual type systems, by contrast, are often designed atop
-an existing dynamically-typed language~\cite{tobin-hochstadt:2016}.
-Many R programs already exist,
-so a successful gradual type system must
-work well with extant code.
+For statically-typed languages, the type system is usually an inseparable part of the language.
+Since they are designed at the same time, the type system precedes any code written in the language.
+Gradual type systems, by contrast, are often designed atop an existing dynamically-typed language~\cite{tobin-hochstadt:2016}.
+Many R programs already exist, so a successful gradual type system must work well with extant code.
+Thus, gradual type system designers must first understand existing language idioms, and then work to accommodate them.
+Our tool assists during the understanding phase of development.
 
-We are not aware of any previous work that applies
-visualization techniques for the purpose of programming language design.
-However, there are existing publications on understanding
-software through visualization.
-Bohnet et al.~\cite{bohnet:2009} visualize massive program execution traces
-to improve developer comprehension of large software systems.
+We are not aware of any previous work that applies visualization techniques for the purpose of programming language design.
+However, there are existing publications on understanding software through visualization.
+Bohnet et al.~\cite{bohnet:2009} visualize massive program execution traces to improve developer comprehension of large software systems.
 Their tool applies pruning and summarization techniques to reduce trace size.
-They focus on call topology,
-describing the structure of function invocations,
-rather than types.
+They focus on call topology, describing the structure of function invocations, rather than types.
 Telea et al.~\cite{telea:2009} also develop a methodology for displaying call structure.
-However, they visualize call graphs of systems in their entirety,
-rather than individual execution traces.
-\typevis has elements of both,
-investigating the system as a whole,
-but doing so via execution traces.
-In addition to standard call graphs,
-LaToza and Myers~\cite{latoza:2011} adorn nodes and edges with extra information,
-allowing the user to gain further understanding of the system
-under examination.
-This includes grouping methods that come from the same class---a
-basic form of type data.
+However, they visualize call graphs of systems in their entirety, rather than individual execution traces.
+\ourmethod has elements of both, investigating the system as a whole, but doing so via execution traces.
+In addition to standard call graphs, LaToza and Myers~\cite{latoza:2011} adorn nodes and edges with extra information, allowing the user to gain further understanding of the system under examination.
+This includes grouping methods that come from the same class --- a basic form of type data.
 
@@
 
-The data set we used comes from
-Turcotte et al.~\cite{turcotte:2020}
-and consists of traces from \pkgNumAnalyzed packages,
-containing over $760,000$ lines of R code
-and $534,000$ lines of native code.
-All libraries were obtained from CRAN,
-the most popular R code repository.
-Packages hosted on CRAN are curated and satisfy
-certain quality standards.
-Out of the over $15,000$ packages
-distributed on CRAN,
-the subset of 412 packages considered here
-all had code coverage above 65\%.
-\todo{Confirm with Alexi this paragraph. Get proper BibTeX entry.}
+We used the data set from Turcotte et al.~\cite{turcotte:2020} that consists of traces from over $760,000$ lines of R code and $534,000$ lines of native code.
+This code was obtained from CRAN, a curated repository of R packages that is widely used in the community.
+The data set consists of all packages with at least $65\%$ code coverage and at least $5$ reverse dependencies (clients using that package).
+Out of the over $15,000$ packages distributed on CRAN, this amounted to \pkgNumAnalyzed packages.
 
-Execution traces were recorded by running a dynamic analysis
-over the test, example, and vignette code of each package.
-A vignette is a form of documentation that weaves
-together prose and executable code.
-An execution trace is a list of function type signatures
-containing, among other attributes, the function name
-and type tags for each argument and return value.
-For example, the function call {\tt !(FALSE)}
-would be recorded as the type signature {\tt logical $\to$ logical},
-meaning that the function {\tt !} took a logical value (i.e. a boolean)
-and returned a logical value.
-The type tags are assigned by the dynamic analysis
-and refine the actual type tags at run-time.
-For instance, the analysis
-assigns the scalar tag {\tt double} to a singleton vector of double values
-because R does not have native support for scalars.
+Execution traces were recorded by running a dynamic analysis over the test, example, and vignette code of each package.
+A vignette is a form of documentation that weaves together prose and executable code.
+An execution trace is a list of function type signatures containing, among other attributes, the function name and type tags for each argument and return value.
+For example, the expression {\tt 0 < 1} would be recorded as the type signature {\tt integer $\to$ integer $\to$ logical}, meaning that the function {\tt <} took two integer values and returned a logical value (i.e.
+a boolean).
+The type tags are assigned by the dynamic analysis and refine the actual type tags at run-time.
+For instance, the analysis assigns the scalar tag {\tt double} to a singleton vector of double values because R does not have native support for scalars.
 The data set is a table of reduced execution traces.
-Because \typevis was designed for the analysis of types,
-not all information in the traces is relevant.
-%that have been processed to reduce some fields.
-%Traces were generated via a dynamic analysis
-%that instruments the R virtual machine with hooks
-%that intercept the required data at run-time.
+Because \ourmethod was designed for the analysis of types, not all information in the traces is relevant.
 
-While \typevis has only been instantiated with
-data from R execution traces, the architecture is not limited to R.
-Any dynamically-typed language, %with the right run-time
-%hooks could generate data for the tool.
-or statically-typed language with run-time tag information,
-could be instrumented to generate data for the tool.
-As long as the trace output adheres to
-the correct schema, \typevis will work properly without any modifications.
+While \ourmethod has only been instantiated with data from R execution traces, the architecture is not limited to R.
+Any dynamically-typed language, or statically-typed language with run-time tag information, could be instrumented to generate data for the tool.
+As long as the trace output adheres to the correct schema, \ourmethod will work properly without any modifications.
 
@@
 
+\begin{figure*}
+ \centering
+ \includegraphics[width=\linewidth]{img/typevis.png}
+ \caption{An overview of \ourmethod, with type flows on the left and the package hierarchy on the right.
+Here, the {\tt length.unit} and {\tt valid.unit} functions from the {\tt grid} package are both selected and displayed simultaneously.}
+ \label{fig:typevis}
+\end{figure*}
+
 \section{Task Abstraction}
 
-We interviewed the primary researcher developing a gradual type system for R
-to inventory relevant domain-specific tasks.
-Additionally, we sought his and other researchers' feedback
-throughout the creation of \typevis
-to ensure it was adequately meeting requirements.
+We interviewed the primary researcher developing a gradual type system for R (and third author of this work) to inventory relevant domain-specific tasks.
+Additionally, we sought his and other researchers' feedback throughout the creation of \ourmethod to ensure it was adequately meeting requirements.
 The domain-specific goals identified were:
@@
 \begin{enumerate}
-\item {\bf Find Function.} Quickly search for a particular function and identify basic information. This includes how often the function is called and which package contains it.
-\item {\bf Determine Types.} Given a specific function,
-  determine its argument and return types, and identify the frequency of specific type signatures.
-  A user should be able to readily infer, for example,
-  if a function is polymorphic (i.e. accepts arguments of multiple different types) or if it is a predicate (i.e. returns a logical value).
-\item {\bf Compare Signatures.} Compare the occurrence of type signatures within a specific function, and across many different functions.
+\item {\bf Find Function.}
+  Quickly search for a particular function and identify basic information.
+  This includes how often the function is called and which package contains it.
+\item {\bf Determine Types.}
+  Given a specific function, determine its argument and return types, and identify the frequency of specific type signatures.
+  A user should be able to readily infer, for example, if a function is polymorphic (i.e. accepts arguments of multiple different types) or if it is a predicate (i.e. returns a logical value).
+\item {\bf Compare Signatures.}
+  Compare the occurrence of type signatures within a specific function, and across many different functions.
   A user must be able to isolate individual signatures and exmaine them in the context of others.
@@
 These goals can be classified using an abstract task analysis framework.
-See table~\ref{tab:tasks} for a categorization of these domain tasks
-with the mid-level and low-level abstract task taxonomy
-established by Brehmer and Munzner~\cite{brehmer:2013}.
+See \cref{tab:tasks} for a categorization of these domain tasks with the mid-level and low-level abstract task taxonomy established by Brehmer and Munzner~\cite{brehmer:2013}.
 
@@
 
-\begin{figure*}
- \centering
- \includegraphics[width=\linewidth]{img/typevis.png}
- \caption{The main view of \typevis, with type flows on the left and the package hierarchy on the right.
-Here, the {\tt length.unit} and {\tt valid.unit} functions from the {\tt grid} package are both selected and displayed simultaneously.}
- \label{fig:typevis}
-\end{figure*}
+\ourmethod is an interactive web-based visualization that permits real-time data exploration, even with large data sets.
+The default data set of R execution traces, described in \cref{sec:data}, contains over a million rows.
 
-\typevis is an interactive web-based visualization
-that permits real-time data exploration,
-even with large data sets.
-The default data set of R execution traces,
-described in \S\ref{sec:data},
-contains over a million rows.
+\Cref{fig:typevis} shows the heart of \ourmethod --- the type flow panel on the left and the function filtering panel on the right.
+Type signatures shown on the left reflect the subset of the data selected on the right.
+By default, if no subset is selected, the type flow panel shows several of the most frequent function type signatures in the entire data set.
+Various aspects of the system, such as the number of type signatures shown at once, can be tweaked by the user if desired; for this, a separate \settingsTab tab is provided.
 
-\figref{fig:typevis} shows
-the heart of \typevis---the type flow panel on the left
-and the function filtering panel on the right.
-Type signatures shown on the left reflect
-the subset of the data selected on the right.
-By default, if no subset is selected, the type flow
-panel shows several of the most frequent function type signatures
-in the entire data set.
-%in its initial configuration.
-Various aspects of the system,
-such as the number of type signatures shown at once,
-can be tweaked by the user if desired;
-for this, a separate \settingsTab tab is provided.
+\subsection{Hierarchical Package Visualization}
 
-\subsection{Function Filtering}
+The right-hand panel of \cref{fig:typevis} contains three elements: a search bar, a package treemap, and a function treemap.
+All three components are linked.
+Selecting a package in the top treemap causes the bottom treemap to show only functions defined in that package; additionally, the type flow panel gets updated with the most frequent function type signatures in the package.
+Selecting a function in the bottom treemap makes the type flow panel show only signatures of that function.
+Entering a function name in the search bar updates the type flow panel immediately.
 
-The right-hand panel of \figref{fig:typevis}
-contains three elements: a search bar,
-a package treemap,
-and a function treemap.
-All three components are linked.
-Selecting a package
-in the top treemap causes the bottom treemap
-to show only functions defined in that package;
-additionally, the type flow panel gets updated
-with the most frequent function type signatures in the package.
-Selecting a function in the bottom treemap
-makes the type flow panel show only signatures
-of that function.
-Entering a function name in the search bar
-updates the type flow panel immediately.
+These different components support at least two modes of search: quick function lookup and package exploration.
+If a user has a specific function in mind, they can type its name into the search bar and see its type signatures straight away; the search bar supports autocomplete that is able to select one function among thousands.
+If, on the other hand, one wants to browse all available packages and functions, the treemap view is more applicable.
+The area of a block in the function treemap (package treemap) conveys the log-scaled frequency of calls to that function (to all functions defined in that package).
+Log-scaling and pagination keep elements of the treemaps legible, even when absolute size differences are significant.
 
-These different components support at least two modes
-of search: quick function lookup and package exploration.
-If a user has a specific function in mind,
-they can type its name into the search bar and see its
-type signatures straight away;
-the search bar supports autocomplete that is
-able to select one function among thousands.
-If, on the other hand, one wants to browse
-all available packages and functions,
-the treemap view is more applicable.
-The area of a block in the function treemap (package treemap)
-conveys the log-scaled
-frequency of calls to that function
-(to all functions defined in that package).
-%Area in the bottom treemap conveys the log-scaled
-%frequency of function calls,
-%and correspondingly the top treemap
-%is scaled based on invocations
-%of functions defined in that package.
-Log-scaling and pagination keep
-elements of the treemaps legible,
-even when absolute size differences are significant.
+The data set can additionally be filtered by analyzed package, in other words, the package containing the call site of a function, not the definition site.
+As mentioned in \cref{sec:data}, the data set contains information from \pkgNumAnalyzed analyzed packages, which are all taken into account by default.
+However, a user can limit the set of analyzed packages, using a search bar in \settingsTab.
 
-The data set can additionally be filtered by analyzed package,
-in other words, the package containing the call site of a function, not the definition site.
-As mentioned in \S\ref{sec:data}, the data set contains information
-from \pkgNumAnalyzed analyzed packages, which are all taken into account
-by default. However, a user can limit the set of analyzed packages,
-using a search bar in \settingsTab.
+Although treemaps are often used for displaying hierarchical data~\cite{shneiderman:1992}, in \ourmethod they have a flat structure.
+The packages in our data are not deeply structured, so the treemaps in \cref{fig:typevis} are only one level deep.
+However, in many programming languages, like Java, package systems contain deeply nested namespace hierarchies.
+The treemap idiom is appropriate for navigation of such hierarchies, and \ourmethod could be modified to handle that.
 
-Although treemaps are often used for displaying
-hierarchical data~\cite{shneiderman:1992},
-in \typevis they have a flat structure.
-The packages in our data are not deeply structured,
-so the treemaps in \figref{fig:typevis} are only one level deep.
-However, in many programming languages, like Java,
-package systems contain deeply nested namespace hierarchies.
-The treemap idiom is appropriate for navigation of such
-hierarchies, and
-%However, if the data set recorded deeply nested packages,
-\typevis could be modified to handle that.
-
-By default, \typevis allows only one function to be selected at a time,
-but it can be configured for multiple function selection.
-In the example from \figref{fig:typevis},
-the {\tt length.unit} and {\tt valid.unit} functions
-from the {\tt grid} package are both selected.
-These functions are rendered simultaneously in the type flow
-panel, allowing the user to make comparisons between them.
-However, if the user wants to explore different functions rather than
-compare them, the default one-function mode is more appropriate.
+By default, \ourmethod allows only one function to be selected at a time, but it can be configured for multiple function selection.
+In the example from \cref{fig:typevis}, the {\tt length.unit} and {\tt valid.unit} functions from the {\tt grid} package are both selected.
+These functions are rendered simultaneously in the type flow panel, allowing the user to make comparisons between them.
+However, if the user wants to explore different functions rather than compare them, the default one-function mode is more appropriate.
 
@@
 
-Prominently featured in the center of \figref{fig:typevis}
-is a parallel sets diagram encoding the type signatures of the selected
-functions as flows.
-Parallel sets,
-proposed by Kosara et al.~\cite{kosara:2006},
-are ideally suited for type signatures.
-Each data point can be thought of as
-a tuple $(f, \tau_1, \ldots, \tau_n, \tau_r)$
-corresponding to the function name,
-argument types,
-and return type.
-Every component of that tuple is categorical,
-and we are interested in visualizing the frequencies of these points.
+Prominently featured in the center of \cref{fig:typevis} is a parallel sets diagram encoding the type signatures of the selected functions as flows.
+Parallel sets, proposed by Kosara et al.~\cite{kosara:2006}, are ideally suited for type signatures.
+Each data point can be thought of as a tuple $(f, \tau_1, \ldots, \tau_n, \tau_r)$ corresponding to the function name, argument types, and return type.
+Every component of that tuple is categorical, and we are interested in visualizing the frequencies of these points.
 This is the exact use-case for parallel sets.
 
-Flows begin at a node labeling a function,
-curve across each argument type in order,
-and terminate at the return type.
-Widths are proportional to how many times a function
-is called with that signature during analysis.
-Flow segments are filled with a hue-varying color scale,
-determined by the type where the segment ends.
-Beneath each node label is an approximate count of either:
-how many times the function was invoked if the node represents a function,
-or how often a value of that type was recorded during analysis if the node represents a type.
+Flows begin at a node labeling a function, curve across each argument type in order, and terminate at the return type.
+Widths are proportional to how many times a function is called with that signature during analysis.
+Flow segments are filled with a hue-varying color scale, determined by the type where the segment ends.
+Beneath each node label is an approximate count of either: how many times the function was invoked if the node represents a function, or how often a value of that type was recorded during analysis if the node represents a type.
 
-In \figref{fig:typevis},
-the {\tt length.unit} function is called with
-two different type signatures at run-time.
-Most frequently, the function is used with the
-signature {\tt double $\to$ integer},
-but a substantial fraction of the time it is
-used as {\tt double[] $\to$ integer}.
-Here, {\tt double[]} stands for an array of
-doubles with arbitrary dimension.
+In \cref{fig:typevis}, the {\tt length.unit} function is called with two different type signatures at run-time.
+Most frequently, the function is used with the signature {\tt double $\to$ integer}, but a substantial fraction of the time it is used as {\tt double[] $\to$ integer}.
+Here, {\tt double[]} stands for an array of doubles with arbitrary dimension.
 
-Interactivity significantly enhances the
-usability of the parallel sets diagram.
-Hovering over a flow will highlight the
-entire path of the flow
-and will fade out all other functions,
-as shown in \figref{fig:typevis}.
-When displaying many flows at the same time,
-highlighting becomes critical
-for user comprehension.
-Additional quantitative information
-about the flow is supplied on-demand while hovering.
-Clicking a flow will focus on that function
-only, filtering away all others;
-double clicking outside flows will bring back the previous view.
-The number of flows shown at a time is limited,
-and will paginate if that threshold is exceeded.
+Interactivity significantly enhances the usability of the parallel sets diagram.
+Hovering over a flow will highlight the entire path of the flow and will fade out all other functions, as shown in \cref{fig:typevis}.
+When displaying many flows at the same time, highlighting becomes critical for user comprehension.
+Additional quantitative information about the flow is supplied on-demand while hovering.
+Clicking a flow will focus on that function only, filtering away all others; double clicking outside flows will bring back the previous view.
+The number of flows shown at a time is limited, and will paginate if that threshold is exceeded.
 
-Flow layout is a necessary consideration
-for a legible visualization~\cite{sugiyama:1981}.
-Figure~\ref{fig:decross} exemplifies what
-happens if flow layout is not adequately computed.
-The top layout was generated without any
-flow decrossing algorithm,
-while the bottom layout applies optimal decrossing.
-Even with highlighting,
-the top visualization is almost impossible to understand
-due to excessive flow overlap.
-To maintain real-time performance,
-we use two different decrossing methods.
-First, the system attempts to construct
-a layout with the globally optimal solution,
-minimizing the amount of flow crossing by
-solving a mixed-integer linear program~\cite{ebner:2005}.
-Unfortunately, for even moderately
-sized graphs this can take too long to compute.
-Therefore, after about $1$ second,
-the optimal algorithm is terminated
-if it has not finished, and the final layout is computed
-by a faster
-algorithm that only minimizes local crossing between layers.
-%computes the layout.
-While the resulting layout might not be ideal,
-it is much better than the one without decrossing at all.
+Flow layout is a necessary consideration for a legible visualization~\cite{sugiyama:1981}.
+\Cref{fig:decross} exemplifies what happens if flow layout is not adequately computed.
+The top layout was generated without any flow decrossing algorithm, while the bottom layout applies optimal decrossing.
+Even with highlighting, the top visualization is almost impossible to understand due to excessive flow overlap.
+To maintain real-time performance, we use two different decrossing methods.
+First, the system attempts to construct a layout with the globally optimal solution, minimizing the amount of flow crossing by solving a mixed-integer linear program~\cite{ebner:2005}.
+Unfortunately, for even moderately sized graphs this can take too long to compute.
+Therefore, after about $1$ second, the optimal algorithm is terminated if it has not finished, and the final layout is computed by a faster algorithm that only minimizes local crossing between layers.
+While the resulting layout might not be ideal, it is much better than the one without decrossing at all.
 
@@
 
-Since the amount of data a user may query is large,
-the application is split into a frontend and a backend.
-The backend is responsible for storing the data
-and fetching required information from the underlying database.
-The frontend serves as an interactive interface to the data
-and provides the visualization.
-This avoids excessive memory consumption and speeds up computations,
-as the client queries data from the server on a need-to-know basis.
+Since the amount of data a user may query is large, the application is split into a frontend and a backend.
+The backend is responsible for storing the data and fetching required information from the underlying database.
+The frontend serves as an interactive interface to the data and provides the visualization.
+This avoids excessive memory consumption and speeds up computations, as the client queries data from the server on a need-to-know basis.
 
-On the frontend, state management, interactivity,
-and DOM manipulation is all done with the Vue framework.
-This differs from many web-based visualizations
-that use D3 for DOM manipulation.
-We chose this setup because \typevis retains many distinct
-components whose state must be kept synchronized.
-This includes not only the visualizations, but also
-user interface elements like the search bar.
-In our experience, Vue's model of reactivity is
-more successful at maintaining consistency across the
-entire application than D3 and vanilla JavaScript.
-While \typevis does not use D3 for DOM manipulation,
-it still does use it for other visualization-related
-tasks. Decrossing is computed using web workers
-to allow early termination if the thread takes too long.
+On the frontend, state management, interactivity, and DOM manipulation is all done with the Vue framework.
+This differs from many web-based visualizations that use D3 for DOM manipulation.
+We chose this setup because \ourmethod contains many distinct components whose state must be kept synchronized.
+This includes not only the visualizations, but also user interface elements like the search bar.
+Vue's model of reactivity can easily maintain consistent state across the entire application.
+While \ourmethod does not use D3 to directly manipulate the DOM, it still does use it for other visualization-related computations.
+Decrossing is calculated using web workers to allow early termination if the thread takes too long.
 
-Our backend is a Node.js application that serves
-JSON with a standard REST-style architecture.
-Data is stored in a SQLite database that has been
-indexed to achieve lookups as fast as possible.
-A stable network connection is required for
-real-time interaction.
+Our backend is a Node.js application that serves JSON with a standard REST-style architecture.
+Data is stored in a SQLite database that has been indexed to achieve lookups as fast as possible.
+A stable network connection is required for real-time interaction.
 
@@
 
+\section{Evaluation}
+
+In accordance with the design study ``lite'' methodology~\cite{syeda:2020}, we conducted a small-scale qualitative usability study and received feedback from $18$ individuals.
+Many respondents had some trouble interpreting the visualization or at least commented that the learning curve was steep.
+Given that the domain is niche and the flow-based idiom for visualizing signatures is novel, this was not unexpected.
+
+To alleviate some of this confusion, we incorporated elements that give explicit explanations for different components.
+These appear as question mark icons that, upon hovering, give a complete description of the corresponding component.
+For more details, we also provide a full page of step-by-step instructions.
+
+%% TODO: Alexi will possibly add evaluation notes here.
+
+%%
+
 \section{Limitations and Future Work}
 
-Designed with a specific use case in mind,
-\typevis suffers from some shortcomings
-and does not fully address the entire range of tasks
-a language designer may need to perform.
+Designed with a specific use case in mind, \ourmethod suffers from some shortcomings and does not fully address the entire range of tasks a language designer may need to perform.
 
-In particular, \typevis does not allow for exploring 
-type information at differing granularities.
-R supports a range of data types
-including ones containing dimensionality information.
-For example, an R array may have type {\tt integer[3]}
-indicating that it is an array of length $3$.
-Every such type will get compressed down %into a single array of
-to a simpler array
-type {\tt integer[]}.
-Reducing the number of types can make the visualization
-more understandable, at the expensive of losing
-precision in type information.
-Ideally, a user would be able to interactively tune
-the level of granularity based on their interest.
-Such a feature would be especially helpful for R,
-because values can come attached with lots of relevant
-data such as the dynamic dispatch method
-or user-specified metadata.
+In particular, \ourmethod does not allow for exploring type information at differing granularities.
+R supports a range of data types including ones containing dimensionality information.
+For example, an R array may have type {\tt integer[3]} indicating that it is an array of length $3$.
+Every such type will get compressed down to a simpler array type {\tt integer[]}.
+Reducing the number of types can make the visualization more understandable, at the expensive of losing precision in type information.
+Ideally, a user would be able to interactively tune the level of granularity based on their interest.
+Such a feature would be especially helpful for R, because values can come attached with lots of relevant data such as the dynamic dispatch method or user-specified metadata.
 
-Filtering is a central mechanism of our system,
-but it also comes at a cost.
-If one has a specific function or several functions
-in mind, then \typevis provides a useful
-local view of that information.
-It does not support, however, any kind of global
-view of type information across a large number of
-functions. When the amount of flows or nodes
-becomes too great, the visualization will start
-paginating, making comparisons across all flows impossible.
-Aggregation and summarization would be key to
-making a global view of the data feasible, but that remains future work.
+Filtering is a central mechanism of our system, but it also comes at a cost.
+If one has a specific function or several functions in mind, then \ourmethod provides a useful local view of that information.
+It does not support, however, any kind of global view of type information across a large number of functions.
+When the amount of flows or nodes becomes too great, the visualization will start paginating, making comparisons across all flows impossible.
+Aggregation and summarization would be key to making a global view of the data feasible, but that remains future work.
 
@@
 
-We have presented \typevis, an interactive visualization
-of run-time type information
-that supports real-time analysis and exploration.
-The tool was instantiated with a massive data set of
-execution traces from a corpus of popular R packages.
+We have presented \ourmethod, an interactive visualization of run-time type information that supports real-time analysis and exploration.
+The tool was instantiated with a massive data set of execution traces from a corpus of popular R packages.
 
-\typevis provides facilities for filtering, searching,
-understanding, and comparing type information.
-Type signatures of function calls
-are represented as flows traversing
-each argument type and return type in sequence.
-Quick function lookup is supported by a search bar,
-with package and function treemaps facilitating
-function exploration.
-Various interactions provide additional details and
-assist users with comprehension and navigation.
+\ourmethod provides facilities for filtering, searching, understanding, and comparing type information.
+Type signatures of function calls are represented as flows traversing each argument type and return type in sequence.
+Quick function lookup is supported by a search bar, with package and function treemaps facilitating function exploration.
+Various interactions provide additional details and assist users with comprehension and navigation.
 
-Our system is aimed at elucidating how R programmers
-use the language in the field.
-Specifically, the data and visualization will be used to
-inform the design of a future gradual type system for R.
-However, \typevis is not tied to the R ecosystem
-and is suitable to use for any %dynamically-typed language.
-language that supports run-time type information.
-The visualization could yield insights
-useful for purposes other than gradual typing,
-and beyond what we have considered.
+Our system is aimed at elucidating how R programmers use the language in the field.
+Specifically, the data and visualization will be used to inform the design of a future gradual type system for R.
+However, \ourmethod is not tied to the R ecosystem and is suitable to use for any language that supports run-time type information.
+The visualization could yield insights useful for purposes other than gradual typing, and beyond what we have considered.
 
-We hope that, in the future, programming language designers
-will collect and use empirical data about their languages
-more extensively.
-Such data could be used to make informed decisions about design,
-where the cost of making a mistake is high.
-Visualization will be unavoidable in these scenarios
-since semantically-meaning information about programs
-is rich in structure.
-Generic, canned visualizations will not be sufficient.
-\typevis is a first step in this direction.
+We hope that, in the future, programming language designers will collect and use empirical data about their languages more extensively.
+Such data could be used to make informed decisions about design, where the cost of making a mistake is high.
+Visualization will be unavoidable in these scenarios since semantically-meaning information about programs is rich in structure.
+Generic, canned visualizations will not be sufficient. \ourmethod is a first step in this direction.
 
 \acknowledgments{
-  The authors wish to thank
-  Alexi Turcotte,
-  Cody Dunne,
-  Sara Di Bartolomeo,
-  Younes El Idrissi Yazami,
-  Peta Maj,
-  the students of CS 7250,
-  and contributors to all the open source projects that \typevis uses.
+  The authors wish to thank Younes El Idrissi Yazami, Petr Maj, the students of CS 7250, and contributors to all the open source projects that \ourmethod uses.
   This work was supported in part by a Northeastern Graduate Fellowship.
@@
 
-\bibliographystyle{abbrv-doi}
+\bibliographystyle{abbrv-doi-hyperref}
 
